### 2024-11-13 Data-Paper/Manuscript-Data-Package.Rmd
Data paper

---
title: "ActiveCA: an open data product on active transportation episodes in Canada"
author:
  - name: "Anon1"
    email: "Anon1@anon.org"
    affiliation: CODE
    footnote: 1
  - name: "Anon2"
    email: "Anon2@anon.org"
    affiliation: CODE
  - name: "Anon3"
    email: "Anon3@anon.org"
    affiliation: CODE
address:
  - code: CODE
    address: "Some school"
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
  "This paper describes the data set of the {ActiveCA} R data package. {ActiveCA} contains open data products to obtain impedance functions for active transportation modes in Canada, retrieved from the General Social Survey collections from 1986 to 2015. The package provides data tables on walking and cycling episodes, detailing the trip origins, destinations, and duration of each episode. The origin and destination categories covers a wide variety of locations, such as home, work or school, libraries, museums, restaurants, bars, sports centers, health clinics, place of worship, and others. Addittionally, the package details the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency. {ActiveCA} enables users to calculate the average travel time for each origin-destination combination for each survey year, considering the two active transportation modes: walking and cycling. For each Census Metropolitan and Agglomeration Area, we estimated distance-decay curves (impedance functions) for each destination, transportation mode and year. The package will continue to expand with contributions from the authors and the broader community through requests in the future. {ActiveCA} is freely accessible for exploration and download from the associated Github repository, where the documentation and code involved in creating and manipulating data and all open data products are detailed."

keywords: Active; mobility; walking; cycling; travel time; impedance; transportation; R 
classoption:
  - Royal
  - times
bibliography: bibfile.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  warning = FALSE,
  comment = '', 
  out.width = "1\\linewidth")
```

```{r install-data-package, include=FALSE}
if (!require("ActiveCA", character.only = TRUE)) {
      remotes::install_github("dias-bruno/ActiveCA",
                        build_vignettes = TRUE)
}

rm(list=ls())
```

```{r load-packages, include=FALSE, cache=FALSE}
library(ActiveCA)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(reshape2)
library(rlang)
library(scales)
library(gridExtra)
library(ggpubr)
library(grid)
library(stats)
```

```{r load-data, include = FALSE}
load(paste0(here(), "/data/gss_episodes.Rda"))
load(paste0(here(), "/data/gss_impedances.Rda"))
```

# Introduction

This paper presents the open data product {ActiveCA}. Open data products
(ODPs) are the outcome of a transparent process that transforms raw data
(open or not) into analysis-ready data, in which all stages of
development follow open principles [@arribas-bel2021]. ODPs differ from
general open data due to their high utility, added value and open
availability. The product presented in this document is an `R` data
package that currently consists of processed data tables retrieved from
the Time Use collections of the General Social Surveys (GSS) from 1986
to 2015 [@statisticscanada2024], aimed to obtain impedance functions for
active transportation modes in Canada.

To create this `R` data package, we collected, cleaned and processed the
Time Use collections from the GSS surveys to make them ready for
analysis. An `R` data package contains code, data and documentation in a
standardized format that can be installed by `R` users via a centralized
software repository, such as CRAN (Comprehensive R Archive Network) and
GitHub. Although the GSS surveys are publicly sourced and managed by
Statistic Canada, preparing them for analysis can be time-consuming,
tedious and perhaps not even possible for those who try, due to a lack
of documentation or prior knowledge.

The aim of this paper is to walk readers through the data sets and
invite others to experiment in its uses and applications. {ActiveCA} is
freely available on GitHub for all to install and freely use in the
spirit of open and reproducible research. Although {ActiveCA} was
designed to obtain impedance functions, we admit and hope that its use
can be adopted in various applications that even go beyond the range of
possibilities we have imagined. Not only the data, but also all the code
documenting the processing methodology is available for consultation and
evaluation in its repository. This package contributes to reducing the
barrier to using the information contained in GSS surveys to provide
data-driven decisions in transportation analysis.

# General Social Survey (GSS) collection

Statistics Canada [-@statisticscanada2024] conducts GSS surveys to
obtain data on social trends to track changes in Canadians' living
conditions and well-being over time. This survey is used to understand
how citizens spend and manage their time and what factors contribute to
their happiness and stress. Created in 1985, the survey is part of a
series of independent, annual, and cross-sectional surveys.

In addition to the main topic, each GSS cycle includes new content that
addresses emerging and policy-relevant issues. Every five to seven
years, the Time Use Surveys [@statisticscanada2022] collect data on
respondents' participation and time spent on a wide range of everyday
activities using a 24-hour retrospective diary, with information on the
location of these activities (e.g. at home, at work, etc.) and, for
non-personal activities, the people who were present with the respondent
at the time of the activity. In addition, time-use surveys also cover
topics related to leisure time, work-life balance, health, commuting,
culture and sports, and many others.

The most recent time use survey was carried out in 2022 [@wray2024].
However, the 2022 dataset has not been fully published and, because of
this, our analysis focused on the surveys from 1986 to 2015 (1986, 1992,
1998, 2005, 2010 and 2015). Time Use surveys are composed of two data
sets, the main one and the episode file, explained in the following
subsections.

## The Main File

The Main File compiles an large array of aggregated data, summarizing
the answers to the questionnaire and derived variables that summarize
the respondents' time use across different activities, locations, and
social interactions. This file documents the time and duration that
respondents allocate to each activity and location. The Main File
provides a overview of daily routines and social dynamics, not focusing
on individual activity episodes. Additionally, this file categorizes
activities into bigger groups and subcategories, facilitating the data's
analytical utility with additional metrics such as total transit time,
duration spent with household members, and counts of activities and
episodes.

## The Episode File

The Episode File records detailed data for each activity episode
reported by respondents. The entry includes the start and end times,
duration, location, and accompanying social context, informing when and
where activities occurred and with whom. The file distinguishes itself
by focusing on individual episodes rather than respondents, with the
data structured around the numerous activity instances that compose a
day of the respondent. Although respondent-specific characteristics are
not included within the Episode File, it is possible to link the Main
File and the Episode File by using an identifiable variable present in
both data sets.

# Descriptive statistics

For each year available from the Time Use surveys, we reviewed the
episode files to identify cases with activities listed as walking or
cycling and selected the activities immediately before and after the
mobility episode. The {ActiveCA} package includes all
`r nrow(gss_episodes)` episodes that recorded walking or cycling as a
mode of transportation, with trip durations ranging from 0 to 900
minutes. Among the analyses possible with this package, Table
\ref{tab:table-01} presents descriptive statistics on walking and
cycling trips between 1986 and 2015, including metrics such as the count
of recorded trips (count), and measures of trip duration in minutes:
maximum (max), mean, median, and minimum (min). The 1986 survey did not
include bicycle trips.

```{r table-01, echo=FALSE, cache=FALSE}

# Calculate the count of trips for each mode and year
# Summarize counts by year and mode
episodes_counts <- gss_episodes %>%
  group_by(YEAR, MODE) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = YEAR, values_from = count)

# Calculate min, mean, and max statistics
stats_data <- gss_episodes %>%
  group_by(YEAR, MODE) %>%
  summarise(
    min = min(DURATION),
    mean = mean(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(min, mean, median, max), names_to = "Statistic", values_to = "Value") %>%
  pivot_wider(names_from = YEAR, values_from = Value)

# Combine counts and stats into one data frame
final_data <- bind_rows(
  episodes_counts %>% mutate(Statistic = "count"),
  stats_data
) %>%
  mutate(MODE = factor(MODE, levels = c("Walking", "Cycling"))) %>%
  arrange(MODE, Statistic) %>%
  mutate(across(where(is.numeric), ~ round(.x, 0)))

# Rename the 'MODE' column to 'Mode' before generating the table
final_data <- final_data %>%
  rename(Mode = MODE) %>%
  dplyr::select(Mode, Statistic, everything())

# Generate the table with 'Statistic' as the second column
styled_ch03_table_01 <- final_data %>%
  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
        caption = "\\label{tab:table-01}Descriptive statistics for episodes with active transport records",  
        align = c('l', 'l', rep('c', ncol(final_data) - 2))) %>%
  kable_styling(full_width = FALSE, font_size = 10) %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c(" " = 2, "Year" = ncol(final_data) - 2)) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major")

# Print the table
styled_ch03_table_01
```

Table \ref{tab:table-01} shows that the median values for walking trips
range between 5 and 10 minutes, while cycling trips have a consistent
median of 15 minutes since 1998. The table also highlights very high
maximum values, particularly for walking trips, with recorded episodes
exceeding 4 hours in all cases.

Table \ref{tab:table-02} and \ref{tab:table-03} provide descriptive
statistics for the two modes of transportation, split by destination
categories, from 1986 to 1998 and from 2005 to 2015, respectively. In
Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
walking trips destined for `home` had the highest medians. However, by
1998, the highest medians shifted to trips to `work or school`, a
transition that also occurred for cycling trips between 1992 and 1998.
Table \ref{tab:table-03} indicates that the median duration for trips to
`home` and `work or school` remained at 10 minutes.

```{r table-02, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}
# Filter for selected years
episodes_stats_before <- gss_episodes %>%
  filter(YEAR < 2005) %>%
  group_by(YEAR, dest_label, MODE) %>%
  summarise(
    count = n(),
    min = min(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    .groups = "drop"
  )

# Calculate percentages
total_counts <- episodes_stats_before %>% 
  group_by(YEAR, MODE) %>% 
  summarise(total_count = sum(count), 
            .groups = "drop")

episodes_stats_before <- left_join(episodes_stats_before, total_counts, by = c("YEAR", "MODE"))

episodes_stats_before$Percentage <- round((episodes_stats_before$count / episodes_stats_before$total_count) * 100, 1)

# Pivot data for the table layout
wide_data <- episodes_stats_before %>%
  pivot_wider(
    id_cols = c(dest_label, MODE),
    names_from = YEAR,
    values_from = c(min, median, max, Percentage)
  )  %>% 
  arrange(dest_label, MODE, desc(Percentage_1998))

wide_data <- wide_data %>%
  group_by(MODE) %>%
  mutate(MODE = if_else(row_number() == 2, as.character(MODE), "")) %>%
  ungroup() %>%
 dplyr::select(
    dest_label, MODE,
    min_1986, median_1986, max_1986, Percentage_1986,
    min_1992, median_1992, max_1992, Percentage_1992,
    min_1998, median_1998, max_1998, Percentage_1998
  )

styled_ch03_table_02 <- wide_data %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(wide_data))),
    caption = "\\label{tab:table-02}Comparison of travel statistics by mode and destination: 1986, 1992, 1998",
    col.names = c("Destination", "Mode*","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
  ) %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 6) %>%
  # column_spec(1, width = "1cm") %>%  # Set width of 'Dest' column to 1.5cm
  # column_spec(2:14, width = "0.5cm") %>%  # Set width of remaining columns to 0.7cm
  add_header_above(c(" " = 2, "1986" = 4, "1992" = 4, "1998" = 4)) %>%
  row_spec(0, bold = TRUE, align = "c") %>%
  column_spec(6, border_right = TRUE) %>%
  column_spec(10, border_right = TRUE) %>% 
  footnote(general = c("* The symbols used in this table represent the following: `Min` denotes the minimum time to reach the destination; `Max` denotes the maximum time to reach the destination; `(%)` indicates a percentage of the total time to reach the destination; `Med` refers to the median time to reach the destination"), threeparttable = T)

styled_ch03_table_02 
```

```{r table-03,echo=FALSE, cache=FALSE, warning=FALSE}
# Rename certain 'dest_label' values
gss_episodes_renamed <- gss_episodes %>%
  mutate(dest_label = case_when(
    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
    dest_label == "Restaurant, bar or club" ~ "Restaurant",
    dest_label == "Sports centre, field or arena" ~ "Sport area",
    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
    dest_label == "In the neighbourhood" ~ "Neighbourhood",
    dest_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ dest_label  # Keep all other values as they are
  ), 
  orig_label = case_when(
    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
    orig_label == "Restaurant, bar or club" ~ "Restaurant",
    orig_label == "Sports centre, field or arena" ~ "Sport area",
    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
    orig_label == "In the neighbourhood" ~ "Neighbourhood",
    orig_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ orig_label))

# Filter for walking and cycling trips, selected years, and duration <= 90
episodes_stats_after <- gss_episodes_renamed %>%
  filter(YEAR > 2000) %>%
  group_by(YEAR, dest_label, MODE) %>%
  summarise(
    count = n(),
    min = min(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    .groups = "drop"
  )

# Calculate percentages
total_counts <- episodes_stats_after %>% group_by(YEAR, MODE) %>% summarise(total_count = sum(count), .groups = "drop")
episodes_stats_after <- left_join(episodes_stats_after, total_counts, by = c("YEAR", "MODE"))
episodes_stats_after$Percentage <- round((episodes_stats_after$count / episodes_stats_after$total_count) * 100, 1)

# Pivot data for the table layout
wide_data <- episodes_stats_after %>%
  pivot_wider(
    id_cols = c(dest_label, MODE),
    names_from = YEAR,
    values_from = c(min, median, max, Percentage)
  ) %>%
 dplyr::select(
    dest_label, MODE,
    min_2005, median_2005, max_2005, Percentage_2005,
    min_2010, median_2010, max_2010, Percentage_2010,
    min_2015, median_2015, max_2015, Percentage_2015
  ) %>% 
  arrange(dest_label, MODE, desc(Percentage_2015)) %>%
  group_by(MODE) %>%
  mutate(MODE = if_else(row_number() == 2, as.character(MODE), "")) %>%
  ungroup()

styled_ch03_table_03 <- wide_data %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(wide_data))),
    caption = "\\label{tab:table-03}Comparison of travel statistics by mode and destination: 2005, 2010, 2015",
    col.names = c("Destination", "Mode*","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
  ) %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                font_size = 6) %>%
  column_spec(1, width = "2cm") %>%  # Set width of 'Dest' column to 1.5cm
  column_spec(2:14, width = "0.4cm") %>%  # Set width of remaining columns to 0.7cm
  add_header_above(c(" " = 2, "2005" = 4, "2010" = 4, "2015" = 4)) %>%
  row_spec(0, bold = TRUE, align = "c") %>%
  column_spec(6, border_right = TRUE) %>%
  column_spec(10, border_right = TRUE) %>% 
  footnote(general = c("* The symbols used in this table represent the following: `Min` denotes the minimum time to reach the destination; `Max` denotes the maximum time to reach the destination; `(%)` indicates a percentage of the total time to reach the destination; `Med` refers to the median time to reach the destination"), threeparttable = T)

styled_ch03_table_03
```

{ActiveCA} also enables visual analysis with traditional exploratory
data analysis techniques. Figures \ref{fig:figure-01} and
\ref{fig:figure-02} show walking and cycling trips from 1992 and 2015
through heat maps. These maps use color gradients to represent the
percentage of trips between various origins and destinations, with
darker colors indicating higher percentages and lighter colors
representing less frequent routes. To avoid overwhelming the reader, we
omitted the heat maps for the other years analyzed.

```{=html}
<!---
I`m not surte if this last phrase makes sense in english! "To avoid overwhelming the reader", "Para nao sobrecarregar o leitor" 
-->
```
In 1992, walking trips with `home` as both the origin and destination
made up the majority, accounting for about 30% of all walking trips.
These trips often involved leisure activities, like short walks or dog
walking. Following this, trips from `home` to `work or school` comprised
18% of walking trips. Overall, `home` emerged as a crucial hub, either
as an origin or destination, with only 5% of trips not involving `home.`
By 2015, `home` remained a significant node, but new locations
distributed the proportion of trips to areas not considered in 1992. In
2015, the highest proportion of trips were from `home` to
`work or school` (12%) and vice versa (11%). `home` to `home` accounted
for 8% of trips, and `grocery stores` became a notable destination for
those leaving `home` (6%), surpassing trips to `other's home` (4%).

```{r heatmaps, echo=FALSE, cache=FALSE, include=FALSE}
gss_heatmap <- gss_episodes_renamed %>%
  mutate(WGHT_EPI = ifelse(is.na(WGHT_EPI), 1, WGHT_EPI)) %>% 
  filter(YEAR %in% c(1992,2015))

for(year in unique(gss_heatmap$YEAR)){

  for(mode in unique(gss_heatmap$MODE)){

    trip_counts <- gss_heatmap %>%
      filter(YEAR == year, MODE == mode) %>% 
      group_by(orig_label, dest_label) %>%
      summarize(count_each_trip = n(), 
                weighted_count = sum(WGHT_EPI), .groups = "drop")

      # Calculate the total weighted count for normalization
      total_weighted_count <- sum(trip_counts$weighted_count)

      # Calculate percentages and round to 2 decimal places
      trip_counts <- trip_counts %>%
        mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

      # Reshape data for heatmap
      heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

      # Convert the data frame to a matrix for the heatmap
      heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
      rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

      # Melt the matrix for use with ggplot
      heatmap_melted <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

      # Define the order of destinations for 2015
      destination_order <- c("Home", "Work or school", "Grocery store", 
                             "Others' home", "Outdoors", "Neighbourhood", 
                             "Cultural venues", "Place of worship", 
                             "Restaurant", "Sport area", "Business")

      # Heatmap for every combination of year and mode
      assign(paste0("heatmap_plot_", year, "_", mode),
               ggplot(heatmap_melted, aes(x = Destination, y = Origin, fill = value)) +
                  geom_tile(color = "white") +
                  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                                       na.value = "gray95", 
                                       limits = c(0, ceiling(max(heatmap_melted$value, na.rm = TRUE))), 
                                       breaks =  c(0, ceiling(max(heatmap_melted$value, na.rm = TRUE))), 
                                       labels = c("0%", paste0(ceiling(max(heatmap_melted$value, na.rm = TRUE)), "%"))) +
                  geom_text(data = subset(heatmap_melted, !is.na(value) & value >= 0), 
                      aes(label = round(value)), 
                      size = 4, color = "black") +
                  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = year) +
                  theme_minimal() +
                  theme(
                  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
                  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
                  axis.title = element_text(size = 16),
                  legend.title = element_text(size = 16, margin = margin(b = 15)), 
                  legend.text = element_text(size = 16),
                  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt")))

             )

  }
} 
```

```{r make-figs, include=FALSE}
heatmap_plot_2015_Walking <- heatmap_plot_2015_Walking +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

heatmap_plot_2015_Cycling <- heatmap_plot_2015_Cycling +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

heatmap_plot_1992_Cycling <- heatmap_plot_1992_Cycling +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

heatmap_plot_1992_Walking <- heatmap_plot_1992_Walking +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

# Combine the plots with adjusted spacing
walking_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Walking, heatmap_plot_2015_Walking, ncol = 2, align = 'hv',  rel_widths = c(1, 1))

# Save the combined plot to a file
ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/walking_hm_fig.jpg"), 
       plot = walking_hm_fig,
       width = 17, 
       height = 8, 
       units = "in", 
       dpi = 300)

# Combine the plots with adjusted spacing
cycling_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Cycling, heatmap_plot_2015_Cycling, ncol = 2, align = 'hv',  rel_widths = c(1, 1))

# Save the combined plot to a file
ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/cycling_hm_fig.jpg"), 
       plot = cycling_hm_fig,
       width = 17, 
       height = 8, 
       units = "in", 
       dpi = 300)
```

```{r figure-01, echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/Data-Paper/Manuscript-figures/walking_hm_fig.jpg"))
```

For cycling trips, Figure \ref{fig:figure-02}, shows that in 1992, when
this mode of transportation was first included as an activity, the
majority of trips were from `home` to `work or school`, accounting for
about 25% of cases. This pattern remained in 2015, with these trips
representing 30% of the cases. However, a notable change occurred in
`home` to `home` trips, which decreased significantly from 19% in 1992
to 5% in 2015.

```{r  figure-02, echo=FALSE, out.width="100%", fig.cap="Percentage of walking Trips Categorized by Origin and Destination", fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/Data-Paper/Manuscript-figures/cycling_hm_fig.jpg"))
```

# Impedance functions for Canadians Metropolitan and Census Agglomerations areas

Impedance functions reveals important information about the travel
behavior of the population, by describing the relationship between the
population at an origin and their likelihood or ability to travel to
specific destinations to access opportunities [@soukhov2024].

Impedance functions are commonly used in accessibility analysis to
calculate the cost of travel between different locations [@hansen1959;
@páez2012; @palacios2022]. However, these functions need to be
calibrated in order to accurately represent travel behavior. One
effective way of calibrating an impedance function is by using the trip
length distribution (TLD) obtained from origin-destination data
[@soukhov2024]. The TLD represents the probability that a certain
proportion of trips will be made at a specific cost, such as travel
time. In this data set, low travel times are associated with a higher
proportion of trips, while high travel times are associated with a lower
proportion of trips.

{ActiveCA} provides calibrated impedance functions for Canadian
Metropolitan and Census Agglomeration areas. For each combination of
year, destination, and transportation mode, we fitted the most suitable
impedance function based on empirical data from the GSS surveys.
{ActiveCA} includes a total of
`r nrow(unique(gss_impedances[,c("YEAR","dest_label","f_name","MODE")]))`
distance-decay functions. These were estimated using the `fitdistrplus`
package [@mullerdutang2015], selecting the distribution with the lowest
Akaike information criterion (AIC) among exponential, gamma, log-normal,
normal, and uniform types.

Table \ref{tab:table-04} shows the best impedance functions for walking
trips to `Work or school` across survey years, where all distributions,
except for a gamma function in 1998, are log-normal. Figure
\ref{fig:figure-03} illustrates these fitted functions (red line)
alongside histograms of empirical travel times (grey bars). As for the
others functions, these examples enable to calculate gravity-based
accessibility measures for active transportation modes across various
destinations and temporal scales in Canadian urban areas.

```{r include = FALSE}
visualize_impedance <- function(df_functions, travel_cost){

  x <- data.frame(t = seq(1, travel_cost, 1))

  for (year in unique(df_functions$YEAR)) {

    name_function <- df_functions[df_functions$YEAR == year, ]$f_name

    subset <- df_functions %>% 
      filter(YEAR == year)

    # Nome da nova coluna com o valor de year
    column_name <- paste0("year_", year)

    # lnorm impedance 
    if (name_function == "lnorm") {
      x <- x %>%
        mutate(!!sym(column_name) := dlnorm(t,
                                             meanlog = subset$est_1[1],
                                             sdlog = subset$est_2[1]))

    # Gamma impedance 
    } else if (name_function == "gamma") {

      x <- x %>%
        mutate(!!sym(column_name) := dgamma(t,
                                             shape = subset$est_1[1],
                                             rate = subset$est_2[1]))
    }else if (name_function == "norm") {  
      x <- x %>%
        mutate(!!sym(column_name) := dnorm(t,
                                             mean = subset$est_1[1],
                                             sd = subset$est_2[1]))
    }else if (name_function == "exp") {  
      x <- x %>%
        mutate(!!sym(column_name) := dexp(t,
                                             exp = subset$est_1[1]))
    }else if (name_function == "unif") {  
      x <- x %>%
        mutate(!!sym(column_name) := dunif(t,
                                             min = subset$est_1[1],
                                             max = subset$est_2[1]))
    }
  }

  return(x)
}
```

```{r , include=FALSE}

max_time <- 100

for(destino in unique(gss_impedances$orig_label)){

  for(mode in unique(gss_impedances[gss_impedances$orig_label == destino, ]$MODE)){

  selected_functions <-  gss_impedances %>% 
  filter(dest_label == destino, MODE == mode) %>%
  dplyr::select("YEAR", "f_name", "est_1", "est_2", "loglik", "aic", "bic") %>%
  mutate(across(c(est_1, est_2), ~ round(., digits = 2)))  %>%
  mutate(across(c(loglik, aic, bic), ~ round(., digits = 0)))  %>%
  unique()

  if(nrow(selected_functions) > 0){
    tld_theoretical <- visualize_impedance(df_functions = selected_functions, travel_cost = max_time)

  # Create an empty list to store the plots
  plots <- list()

# Loop through each unique year

  for(year in unique(gss_impedances$YEAR)){

    # Subset empirical and theoretical data for the current year
    empirical_subset <- gss_impedances %>% 
      filter(dest_label == destino, MODE == mode, YEAR == year, DURATION <= max_time)

    if(nrow(empirical_subset) > 1 ){

      print(paste0(mode," - ", destino," - ", year))

      theoretical_subset <- tld_theoretical %>%
      dplyr::select(t, matches(as.character(year)))

      # Create the ggplot
      plots[[as.character(year)]] <- ggplot(empirical_subset, aes(x = DURATION)) +
        geom_histogram(aes(y = ..density..), binwidth = 5, fill = "#dce1e3", color = "#adb1b3") +
        geom_line(data = theoretical_subset, aes(x = t, y = theoretical_subset[[2]]), color = "#b52640", size = 0.80) +
        labs(title = paste(year)) +
        theme_minimal() +
        scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
        theme(axis.title.x = element_blank(),
            axis.title.y = element_blank(), 
            plot.title = element_text(size = 10, face="bold"))

  }

# Use gridExtra to arrange the plots in a 2x3 layout and add overall axis labels

    if(length(plots) == 1){
      combined_plot <- grid.arrange(
      grobs = plots,
      ncol = 1,
      bottom = text_grob(paste0("Travel time to `", destino ,"` (in minutes) considering `", mode,"` as transportation mode.")),
      left = text_grob("Density", rot = 90)
    )
      ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
       plot = combined_plot,
       width = 9, 
       height = 4, 
       units = "in", 
       dpi = 300)

    } else if(length(plots) == 3){
      combined_plot <- grid.arrange(
      grobs = plots,
      ncol = 3,
      bottom = text_grob(paste0("Travel time to `", destino ,"` (in minutes) considering `", mode,"` as transportation mode.")),
      left = text_grob("Density", rot = 90)
    )
      ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
       plot = combined_plot,
       width = 9, 
       height = 4, 
       units = "in", 
       dpi = 300)
    }else {
      combined_plot <- grid.arrange(
      grobs = plots,
      ncol = 3,
      bottom = textGrob(paste0("Travel time to `", destino ,"` (in minutes) considering `", mode,"` as transportation mode."), gp = gpar(fontsize = 16)),
      left = textGrob("Density", rot = 90, gp = gpar(fontsize = 16))
    )
      ggsave(file = paste0(here::here(),"/Data-Paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"), 
       plot = combined_plot,
       width = 14, 
       height = 8, 
       units = "in", 
       dpi = 300)
    }
  }
}
}
} 
```

```{r table-04, echo=FALSE, cache=FALSE, warning=FALSE}
mode <- "Walking" 
destino <- "Work or school"

gss_impedances %>% 
filter(dest_label == destino, MODE == mode) %>%
  dplyr::select("YEAR","f_name", "est_1","est_2","aic") %>%
  mutate(across(c(est_1, est_2), ~ round(., digits = 2)))  %>%
  mutate(across(c(aic), ~ round(., digits = 0)))  %>%
  unique() %>% 
  kable(format = "latex",
        booktabs = TRUE,
        caption = paste0("\\label{table-04}Impedance functions and AIC for `", mode ,"` trips considering `", destino,"` as destination."), 
        col.names = c("Year", "Impedance function*","Parameter 1*","Parameter 2*", "AIC")) %>% 
    kable_styling(full_width = FALSE, 
                latex_options=c("scale_down"),
                font_size = 10) %>% 
  row_spec(0, bold = TRUE, align = "c") %>% 
  footnote(general = c("* `lnorm` refers to the log-normal distribution. For `lnorm` distributions, `Parameter 1` and `Parameter 2` refer to the mean and standard deviation of the distribution on the logarithmic scaler, respectively. For the `gamma` distribution, `Parameter 1` and `Parameter 2` refer to the rate and shape of the distribution, respectively. `AIC` means Akaike information criterion."), threeparttable = T)
```

```{r figure-03, echo=FALSE, out.width="100%", fig.cap=paste0("Percentage of walking Trips Categorized by Origin and Destination", mode), fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/Data-Paper/Manuscript-figures/impf_",destino,"_",mode,".jpg"))
```

# Concluding remarks

This article introduces {ActiveCA}, an open data product in the form of
an R data package, developed after the collection, cleaning, and
processing of General Social Survey data ranging from 1986 to 2015. The
package provides analysis-ready data on active transportation episodes,
focusing on walking and cycling activities, with information on trip
origins, destinations, and duration. Additionally, the package includes
a series of impedance functions calibrated for various destinations,
considering the different transportation modes and time periods.

The value of {ActiveCA} lies in its transparency, accessibility, and
open infrastructure, which facilitates the addition of complementary
data sets in the future. R users can seamlessly explore GSS walking and
cycling episodes along with calibrated impedance functions, with the
option to suggest enhancements to the package as needed. This article
adopts the structure proposed by Anastasia and Páez [-@soukhov2023],
whose work provided essential guidance for the creation of this package.
Similarly, we aim to contribute to the academic community by promoting
transparent research practices that encourage replication and innovation
in related fields. We believe that {ActiveCA} will serve as a basis for
further research on GSS and for the integration of additional data by
the authors or the wider open source community.

# Declaration of Conflicting Interests

The author(s) declared no potential conflicts of interest with respect
to the research, authorship, and/or publication of this article.

# Funding

The author(s) disclosed receipt of the following financial support for
the research, authorship, and/or publication of this article: This work
was supported by the Social Sciences and Humanities Research Council of
Canada (*More description about the funding source after the review
process*).

# ORCID

Author 1

Author 2

Author 3

# Data availability statement

The {ActiceCA} R data package can be found and installed on Github
(*link*).

# References {#references .unnumbered}