### 2024-11-13 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ActiveCA

{ActiveCA} is an open data product in the form of
an R data package, developed after the collection, cleaning, and
processing of [General Social Survey
(GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221) data ranging from 1986 to 2015. The main objective of the {ActiveCA} package is to identify the most
appropriate impedance functions for active transportation modes for
various destinations and time periods in Canada.  The package also provides analysis-ready data on active transportation episodes, focusing on walking and cycling activities, with information on trip origins, destinations, and duration.

## Installation

You can install the development version of ActiveCA from [GitHub](https://github.com/) with:
``` r
# install.packages("pak")
pak::pak("dias-bruno/ActiveCA")
```