### 2024-12-04 paper/Manuscript-Data-Package.Rmd
Started editing the paper after Bruno's work

---
title: "ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel"
author:
  - name: "Anon1"
    email: "Anon1@anon.org"
    affiliation: CODE
    footnote: 1
  - name: "Anon2"
    email: "Anon2@anon.org"
    affiliation: CODE
  - name: "Anon3"
    email: "Anon3@anon.org"
    affiliation: CODE
address:
  - code: CODE
    address: "Some school"
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
  "This paper describes {ActiveCA}, and open data product with Canadian time use 
data. {ActiveCA} is an `R` data package that contains analysis-ready data related 
to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992),
12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey.
Active travel is characterized by mode, with walking being part of every cycle 
and bicycling starting in 1992. The attributes of active trips are the types of 
locations of origins and destinations, the duration of trips, and episode weights
for expanding the trips to population-wide estimates. Based on the year of the 
survey, a variety of locations are coded. In earlier cycles, these include home,
work or school, and other's home, whereas in later cycles these are augmented with
locations such as grocery stores, restaurants, outdoor destinations, and others.
The geographical resolution includes the province and whether the episode was in
an urban or rural setting."

keywords: Active; mobility; walking; cycling; travel time; time-use; General Social Survey 
classoption:
  - Royal
  - times
bibliography: bibfile.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
editor_options: 
  markdown: 
    wrap: 72
---

<!-- 
FROM THE ABSTRACT:

The package also details the respondent's region characteristics, specifying wheter they live in a metropolitan area and their province of residency. {ActiveCA} also provides pre-estimated impedance functions for active travel for Canadian Metropolitan Area and Census Agglomerations, which can be used in accessibility analysis to calculate the cost of travel between different locations. The package will continue to expand with contributions from the authors and the broader community through requests in the future. {ActiveCA} is freely accessible for exploration and download from the associated Github repository, where the documentation and code involved in creating and manipulating data are detailed.
-->

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  warning = FALSE,
  comment = '', 
  out.width = "1\\linewidth")

opts <- options(knitr.kable.NA = "")
```

```{r install-data-package, include=FALSE}
if (!require("ActiveCA", character.only = TRUE)) {
      remotes::install_github("dias-bruno/ActiveCA",
                        build_vignettes = TRUE)
}

rm(list=ls())
```

```{r load-packages, include=FALSE, cache=FALSE}
# Load packages
library(ActiveCA)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(reshape2)
library(rlang)
library(scales)
library(gridExtra)
library(ggpubr)
library(grid)
library(stats)
```

```{r load-data, include = FALSE}
# Load data
data("gss_episodes")
data("gss_impedances")
```

```{r rename-labels, include = FALSE}
# Renaming 'dest_label' and 'orig_label' values to reduce table space
gss_episodes_renamed <- gss_episodes %>%
  mutate(dest_label = case_when(
    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
    dest_label == "Restaurant, bar or club" ~ "Restaurant",
    dest_label == "Sports centre, field or arena" ~ "Sport area",
    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
    dest_label == "In the neighbourhood" ~ "Neighbourhood",
    dest_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ dest_label  # Keep all other values as they are
  ), 
  orig_label = case_when(
    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
    orig_label == "Restaurant, bar or club" ~ "Restaurant",
    orig_label == "Sports centre, field or arena" ~ "Sport area",
    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
    orig_label == "In the neighbourhood" ~ "Neighbourhood",
    orig_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ orig_label))

gss_impedances_renamed <- gss_impedances %>%
  mutate(dest_label = case_when(
    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
    dest_label == "Restaurant, bar or club" ~ "Restaurant",
    dest_label == "Sports centre, field or arena" ~ "Sport area",
    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
    dest_label == "In the neighbourhood" ~ "Neighbourhood",
    dest_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ dest_label  # Keep all other values as they are
  ), 
  orig_label = case_when(
    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
    orig_label == "Restaurant, bar or club" ~ "Restaurant",
    orig_label == "Sports centre, field or arena" ~ "Sport area",
    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
    orig_label == "In the neighbourhood" ~ "Neighbourhood",
    orig_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ orig_label))
```

# Introduction

The objective of this paper is to introduce {ActiveCA}, an open data product with
time use data from Canada's General Social Surveys. Open data products
(ODPs) are the outcome of a process that transforms raw data (open or not) into 
analysis-ready data, following a transparent process in which all stages of
development follow open principles [@arribas-bel2021]. ODPs, while still open, 
differ from general open data in their degree of ease of access, their heightened 
usability, and potentially the value they add to the raw data.

{ActiveCA} provides analysis-ready data concerning active travel in Canada spanning
a period of almost 40 years. The source of these data is Canada's program of 
General Social Surveys (GSSs). This program is designed to provide cross-sectional
data on topics of interest to improve the well-being of Canadians. As part of this
program, every five to seven years the survey is done on the topic of time use. 
Concretely, {ActiveCA} covers Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005),
24 (2010), and 29 (2015) of the GSS. Time use data in these surveys is coded using
a very fine grain, from time spent in chores, leisure, and sleeping, to time spent
working or at school. These surveys have proved valuable in investigations of 
mobility and quality of life [@spinneyTransport2009], the relationship between
active travel and transit use [@lachapelleLonger2016], and travel behavior and time
poverty [@kimFacing2024], to name but a few examples.

For {ActiveCA}, data we extracted all data needed to characterize active travel 
in Canada, namely, episodes where the activity was identified as moving between 
an origin and a destination, either by walking or cycling. Although Statistics 
Canada offers Public Use Microdata Files and documentation for the GSS program 
[see @statisticscanada2024], accessing these files, and preparing  them for analysis
is not a straightforward matter, given their size and complexity.  The  process 
of extracting information of interest from the source files is  time-consuming, 
tedious, and challenging and/or prone to error due to the expertise required to 
work with these files. To create {ActiveCA} we collected, cleaned, and processed
the cycles of the GSS  surveys concerning time use to make them ready for analysis.

{ActiveCA} is distributed as an `R` package with a number of data 
objects and their documentation. `R` packages contain code, data, and documentation
in a standardized format that can be installed by `R` users via a software repository,
such as CRAN (Comprehensive R Archive Network) or GitHub, which makes them an 
adroit medium to distribute analysis-ready data.

Given the level of interest in active travel [e.g., @mccurdySupport2023], reducing
the barriers to using data contained in rich, but difficult to access and use surveys,
such as Canada's GSSs, is a worthy endeavour that can only improve data-driven 
decisions in transportation, urban, and health policy.

<!-- 
is an `R` data package that provides analysis-ready data from the General Social Survey of Canada  [@statisticscanada2024] regarding active travel. These data can be used to study walking and cycling in Canada for various purposes over time, and to calculate impedance functions for accessibility analysis. The package provides pre-estimated impedance functions for cycling and walking trips for Census Agglomerations and Census Metropolitan Areas.
-->

The rest of this paper discusses the sources of data, and the process implemented
to retrieve and package them. Then, we explain how to use the package and show
some selected examples of analysis to whet the imagination of potential users.
This ODP provides not only data that are easy to use, but also all the code and 
documentation that make this a reproducible research project. In summary, {ActiveCA}
aims to implement and inspire the best principles of open spatial sciences 
[@paez_open_2021; @brunsdon_opening_2021].

# General Social Survey (GSS) collection

Statistics Canada [-@statisticscanada2024] conducts GSS surveys to obtain data 
on social trends to track changes in Canadians' living conditions and well-being
over time. The series of survey on time use are used to understand how Canadian
residents spend and manage their time, and what factors contribute to their 
happiness and stress. The GSS program was created in 1985, and is serialized to 
provide a collection of annual, representative cross-sectional surveys.

The topics of the survey cycle every few years to cover topics that include family,
health, social identity, and every five to seven years time use. The first Canadian
time use survey done as part of the GSS program was conducted in 1986, and the 
most recent was completed in 2015. These Time Use Surveys [@statisticscanada2022]
collect data on respondents' participation and time spent on a wide range of 
everyday activities using a 24-hour retrospective diary, with information on the 
location of these activities (e.g. at home, at work, etc.) and, for non-personal
activities, the people who were present with the respondent at the time of the 
activity. In addition, time-use surveys also cover topics related to leisure time,
work-life balance, health, commuting, culture and sports, and many others.

The Public Use Microdata are released by Statistics Canada in two files: a 
_Main File_ and an _Episode File_. The files are linked by keys that identify 
households, individuals, and episodes (i.e., activities) conducted by individuals.
We discuss these files in more detail in the following section.

## The Main File

The main file of the time use survey compiles a large array of aggregated data, 
summarizing the answers to the questionnaire that describe households and 
individuals, as well as derived variables that summarize the respondents' use of 
time use across different activities, locations, and social interactions. This 
file documents the time and duration that respondents allocate to each activity 
and location. The Main File provides a overview of daily routines and social 
dynamics, not focusing on individual activity episodes. Additionally, this file 
categorizes activities into bigger groups and subcategories, facilitating the 
data's analytical utility with additional metrics such as total transit time, time 
spent with household members, and counts of activities and episodes.

<!-- 
Maybe here provide an example of the structure of this table?
-->

## The Episode File

The episode is a much bigger file that records detailed data for each activity 
episode reported by respondents. Each episode represents a single activity and its
duration, and the sum of all episodes throughout the day adds up to 24 hours. Each
entry in this file includes the start and end times of the activity, the duration,
location, and accompanying social context, informing when and where activities 
occurred and with whom. The focus of the Episode File is not on the characteristics
of the respondents, but on the characteristics of the activities, and the data 
are structured around the numerous activity instances that compose a day of the 
respondent. Although respondent-specific characteristics are not included within
the episode file, it is possible to link the main file and the episode file by 
using a key present in both the Main and Episode Files.

<!-- 
Maybe here provide an example of the structure of this table?
-->

# Methods

For each selected cycle of the GSS surveys, we reviewed the episode files to 
identify episodes of movement that involved walking or cycling. This allowed us 
to select the activities immediately before and after the movement episode. After
that, we labeled the code variables with their appropriate descriptions, identifying
each origin and destination, mode of travel, as well the province and urban 
classification of the episode.

<!--

Show a graphical example of this process: how were active trips identified? 
What was the purpose of finding the preceding and succeeding activities?

-->

<!--

Not needed in this paper:

Additionally, for Census Agglomerations and Census Metropolitan Areas we pre-estimated impedance functions for cycling and walking trips. Impedance functions reveals important information about the travel behavior of the population, by describing the relationship between the population at an origin and their likelihood or ability to travel to specific destinations to access opportunities [@soukhov2024].

Impedance functions are commonly used in accessibility analysis to calculate the cost of travel between different locations [@hansen1959; @páez2012; @palacios2022]. However, these functions need to be calibrated in order to accurately represent travel behavior. One effective way of calibrating an impedance function is by using the trip length distribution (TLD) obtained from origin-destination data [@soukhov2024]. The TLD represents the probability that a certain proportion of trips will be made at a specific cost, such as travel time. In this data set, low travel times are associated with a higher proportion of trips, while high travel times are associated with a lower proportion of trips.

For each combination of year, destination, and transportation mode (walking or cycling), we fitted the most suitable impedance function based on empirical data from the GSS surveys. We used the `fitdistrplus` package [@mullerdutang2015] to estimate the functions, selecting the distribution with the lowest Akaike information criterion (AIC) among exponential, gamma, log-normal, normal, and uniform types.

-->

# How to use {ActiveCA}

<!--

The paper does not really need to report "Results". There are too many potential use cases and we would rather let users find their own results depending on what they are interested in. Instead of results, a section on "How to use the package" is appropriate. There are a number of points to highlight here, for example, how to load the package, call the data sets, in particular the use of episode weights to make sure that any statistical analysis is valid.

-->

## Descriptive statistics

Considering GSS Cycles analyzed, we identified `r nrow(gss_episodes)` episodes that recorded active travel episodes, with trip duration ranging from 0 to 900 minutes, to twelve different destinations. `ActiveCA` includes all these episodes ready for analysis. Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2015, including metrics such as the count
of recorded trips (count), and measures of trip duration in minutes:
maximum (max), mean, median, and minimum (min). The 1986 survey did not
include bicycle trips.

```{r table-01, echo=FALSE, cache=FALSE}
# Calculate the count of trips for each mode and year
episodes_counts <- gss_episodes %>%
  group_by(YEAR, MODE) %>%
  summarise(Count = n(), .groups = "drop") %>%
  pivot_wider(names_from = YEAR, values_from = Count)

# Calculate min, mean, and max statictis (duration)
stats_data <- gss_episodes %>%
  group_by(YEAR, MODE) %>%
  summarise(
    min = min(DURATION),
    mean = mean(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    std = sd(DURATION),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(min, mean, median, max, std), names_to = "Statistic", values_to = "Value") %>%
  pivot_wider(names_from = YEAR, values_from = Value)

# Combine counts and stats into one data frame
final_data <- bind_rows(
  episodes_counts %>% mutate(Statistic = "Count"),
  stats_data
) %>%
  mutate(MODE = factor(MODE, levels = c("Walking", "Cycling"))) %>%
  arrange(MODE, Statistic) %>%
  mutate(across(where(is.numeric), ~ round(.x, 0)))

# Rename the 'MODE' column to 'Mode' before generating the table
final_data <- final_data%>% 
  arrange(MODE) %>%
  rename(Mode = MODE) %>%
  dplyr::select(Mode, Statistic, everything()) %>% 
   mutate(Statistic = case_when(
    Statistic == "max" ~ "Maximum",
    Statistic == "min" ~ "Minimum",
    Statistic == "mean" ~ "Mean",
    Statistic == "median" ~ "Median",
    Statistic == "std" ~ "Standard deviation",
    TRUE ~ Statistic  # Keep all other values as they are
  ))

# Creating a latex table
styled_ch03_table_01 <- final_data %>%
  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
        caption = "\\label{tab:table-01}Descriptive statistics for episodes with active transport records",  
        align = c('l', 'l', rep('c', ncol(final_data) - 2))) %>%
  kable_styling(full_width = FALSE, font_size = 10) %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c(" " = 2, "Year" = ncol(final_data) - 2)) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major")

# Print the table
styled_ch03_table_01
```

Table \ref{tab:table-01} shows that the median values for walking trips
range between 5 and 10 minutes, while cycling trips have a consistent
median of 15 minutes since 1998. The table also highlights very high
maximum values, particularly for walking trips, with recorded episodes
exceeding 4 hours in all cases.

Table \ref{tab:table-02} and \ref{tab:table-03} provide descriptive
statistics for the two modes of transportation, split by destination
categories, from 1986 to 1998 and from 2005 to 2015, respectively. In
Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
walking trips destined for `home` had the highest medians. However, by
1998, the highest medians shifted to trips to `work or school`, a
transition that also occurred for cycling trips between 1992 and 1998.
Table \ref{tab:table-03} indicates that the median duration for trips to
`home` and `work or school` remained at 10 minutes.

```{r table-02, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}
# Filter for selected years and calculate statistics of trip duration
episodes_stats_before <- gss_episodes %>%
  filter(YEAR < 2005) %>%
  group_by(YEAR, dest_label, MODE) %>%
  summarise(
    count = n(),
    min = min(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    .groups = "drop"
  )

# Calculate percentages of trips by mode/year 
total_counts <- episodes_stats_before %>% 
  group_by(YEAR, MODE) %>% 
  summarise(total_count = sum(count), 
            .groups = "drop")

episodes_stats_before <- left_join(episodes_stats_before, total_counts, by = c("YEAR", "MODE"))

episodes_stats_before$Percentage <- round((episodes_stats_before$count / episodes_stats_before$total_count) * 100, 1)

# Pivot data for the table layout
wide_data <- episodes_stats_before %>%
  pivot_wider(
    id_cols = c(dest_label, MODE),
    names_from = YEAR,
    values_from = c(min, median, max, Percentage)
  )  %>% 
  arrange(MODE, dest_label, desc(Percentage_1998))

wide_data <- wide_data %>%
 dplyr::select(
    MODE, dest_label,
    min_1986, median_1986, max_1986, Percentage_1986,
    min_1992, median_1992, max_1992, Percentage_1992,
    min_1998, median_1998, max_1998, Percentage_1998
  ) %>% 
  arrange(MODE, dest_label)

# Latex table
styled_ch03_table_02 <- wide_data %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(wide_data))),
    caption = "\\label{tab:table-02}Comparison of travel statistics by mode and destination: 1986, 1992, 1998",
    col.names = c("Destination", "Mode*","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
  ) %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 6) %>%
  # column_spec(1, width = "1cm") %>%  # Set width of 'Dest' column to 1.5cm
  # column_spec(2:14, width = "0.5cm") %>%  # Set width of remaining columns to 0.7cm
  add_header_above(c(" " = 2, "1986" = 4, "1992" = 4, "1998" = 4)) %>%
  row_spec(0, bold = TRUE, align = "c") %>%
  column_spec(6, border_right = TRUE) %>%
  column_spec(10, border_right = TRUE) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>% 
  footnote(general = c("* The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)

styled_ch03_table_02 
```

```{r table-03,echo=FALSE, cache=FALSE, warning=FALSE}
# Selecting surveys after 2000 and calculating trip duration statistics 
episodes_stats_after <- gss_episodes_renamed %>%
  filter(YEAR > 2000) %>%
  group_by(YEAR, dest_label, MODE) %>%
  summarise(
    count = n(),
    min = min(DURATION),
    median = median(DURATION),
    max = max(DURATION),
    .groups = "drop"
  )

# Calculate percentages of trips by mode/year
total_counts <- episodes_stats_after %>% group_by(YEAR, MODE) %>% summarise(total_count = sum(count), .groups = "drop")
episodes_stats_after <- left_join(episodes_stats_after, total_counts, by = c("YEAR", "MODE"))
episodes_stats_after$Percentage <- round((episodes_stats_after$count / episodes_stats_after$total_count) * 100, 1)

# Pivot data for the table layout
wide_data <- episodes_stats_after %>%
  pivot_wider(
    id_cols = c(dest_label, MODE),
    names_from = YEAR,
    values_from = c(min, median, max, Percentage)
  ) %>%
 dplyr::select(
    MODE, dest_label,
    min_2005, median_2005, max_2005, Percentage_2005,
    min_2010, median_2010, max_2010, Percentage_2010,
    min_2015, median_2015, max_2015, Percentage_2015
  ) %>% 
  arrange(MODE, dest_label)

# Latex table
styled_ch03_table_03 <- wide_data %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(wide_data))),
    caption = "\\label{tab:table-03}Comparison of travel statistics by mode and destination: 2005, 2010, 2015",
    col.names = c("Destination", "Mode","Min*","Med*","Max*", "(%)*", "Min","Med","Max", "(%)", "Min","Med","Max", "(%)")
  ) %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                font_size = 6) %>%
  #column_spec(1, width = "2cm") %>%  # Set width of 'Dest' column to 1.5cm
  #column_spec(2:14, width = "0.4cm") %>%  # Set width of remaining columns to 0.7cm
  add_header_above(c(" " = 2, "2005" = 4, "2010" = 4, "2015" = 4)) %>%
  row_spec(0, bold = TRUE, align = "c") %>%
  column_spec(6, border_right = TRUE) %>%
  column_spec(10, border_right = TRUE) %>% 
  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>% 
  footnote(general = c("* The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination"), threeparttable = T)

styled_ch03_table_03
```

{ActiveCA} also enables visual analysis of active travel in Canada using traditional 
exploratory data analysis techniques. Figures \ref{fig:figure-01} and \ref{fig:figure-02}
show walking and cycling trips from 1992 and 2015 through heat maps. These maps 
use color gradients to represent the percentage of trips between various origins
and destinations, with darker colors indicating higher percentages and lighter 
colors representing less frequent routes. For conciseness, we omitted the heat maps
for the other years analyzed.

```{=html}
<!---
I`m not surte if this last phrase makes sense in english! "To avoid overwhelming the reader", "Para nao sobrecarregar o leitor" 
-->
```

In 1992, walking trips with `home` as both the origin and destination
made up the majority, accounting for about 30% of all walking trips.
These trips often involved leisure activities, like short walks or dog
walking. Following this, trips from `home` to `work or school` comprised
18% of walking trips. Overall, `home` emerged as a crucial hub, either
as an origin or destination, with only 5% of trips not involving `home.`
By 2015, `home` remained a significant node, but new locations
distributed the proportion of trips to areas not considered in 1992. In
2015, the highest proportion of trips were from `home` to
`work or school` (12%) and vice versa (11%). `home` to `home` accounted
for 8% of trips, and `grocery stores` became a notable destination for
those leaving `home` (6%), surpassing trips to `other's home` (4%).

```{r heatmaps, echo=FALSE, cache=FALSE, include=FALSE}

# Creating heatmaps for the years of 1992 and 2015
gss_heatmap <- gss_episodes_renamed %>%
  mutate(WGHT_EPI = ifelse(is.na(WGHT_EPI), 1, WGHT_EPI)) %>% 
  filter(YEAR %in% c(1992,2015))

for(year in unique(gss_heatmap$YEAR)){

  for(mode in unique(gss_heatmap$MODE)){

    trip_counts <- gss_heatmap %>%
      filter(YEAR == year, MODE == mode) %>% 
      group_by(orig_label, dest_label) %>%
      summarize(count_each_trip = n(), 
                weighted_count = sum(WGHT_EPI), .groups = "drop")

      # Calculate the total weighted count
      total_weighted_count <- sum(trip_counts$weighted_count)

      # Calculate percentages and round to 2 decimal places
      trip_counts <- trip_counts %>%
        mutate(percentage = round((weighted_count / total_weighted_count) * 100, 2))

      # Reshape data for heatmap
      heatmap_data <- dcast(trip_counts, orig_label ~ dest_label, value.var = "percentage")

      # Convert the data frame to a matrix for the heatmap
      heatmap_matrix <- as.matrix(heatmap_data[,-1]) # Remove the first column (orig)
      rownames(heatmap_matrix) <- heatmap_data$orig # Assign row names from the 'orig' column

      # Melt the matrix for use with ggplot
      heatmap_melted <- melt(heatmap_matrix, varnames = c("Origin", "Destination"))

      # Define the order of destinations 
      destination_order <- c("Home", "Work or school", "Grocery store", 
                             "Others' home", "Outdoors", "Neighbourhood", 
                             "Cultural venues", "Place of worship", 
                             "Restaurant", "Sport area", "Business")

      # Heatmap for every combination of year and mode
      assign(paste0("heatmap_plot_", year, "_", mode),
               ggplot(heatmap_melted, aes(x = Destination, y = Origin, fill = value)) +
                  geom_tile(color = "white") +
                  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd"), 
                                       na.value = "gray95", 
                                       limits = c(0, ceiling(max(heatmap_melted$value, na.rm = TRUE))), 
                                       breaks =  c(0, ceiling(max(heatmap_melted$value, na.rm = TRUE))), 
                                       labels = c("0%", paste0(ceiling(max(heatmap_melted$value, na.rm = TRUE)), "%"))) +
                  #geom_text(data = subset(heatmap_melted, !is.na(value) & value >= 0), 
                  #    aes(label = round(value)), 
                  #    size = 4, color = "black") +
                  labs(x = "Destination", y = "Origin", fill = "Percentage", caption = year) +
                  theme_minimal() +
                  theme(
                  axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
                  axis.text.y = element_text(size = 16, margin = margin(l = 15)),
                  axis.title = element_text(size = 16),
                  legend.title = element_text(size = 16, margin = margin(b = 15)), 
                  legend.text = element_text(size = 16),
                  plot.caption = element_text(hjust = 0.5, size = 20, face = "bold", margin = margin(t = 28.35, unit = "pt")))

             ) 
  }
} 
```

```{r make-figs, include=FALSE}
# Fixing the size of the heatmaps 
heatmap_plot_2015_Walking <- heatmap_plot_2015_Walking +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

heatmap_plot_2015_Cycling <- heatmap_plot_2015_Cycling +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

heatmap_plot_1992_Cycling <- heatmap_plot_1992_Cycling +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

heatmap_plot_1992_Walking <- heatmap_plot_1992_Walking +
  coord_fixed(ratio = 1) +
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"),  
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14))

# Combine the plots with adjusted spacing
walking_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Walking, heatmap_plot_2015_Walking, ncol = 2, align = 'hv',  rel_widths = c(1, 1))

# Save the combined plot to a file
ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/walking_hm_fig.jpg"), 
       plot = walking_hm_fig,
       width = 17, 
       height = 8, 
       units = "in", 
       dpi = 300)

# Combine the plots with adjusted spacing
cycling_hm_fig <- cowplot::plot_grid(heatmap_plot_1992_Cycling, heatmap_plot_2015_Cycling, ncol = 2, align = 'hv',  rel_widths = c(1, 1))

# Save the combined plot to a file
ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/cycling_hm_fig.jpg"), 
       plot = cycling_hm_fig,
       width = 17, 
       height = 8, 
       units = "in", 
       dpi = 300)
```

```{r figure-01, echo=FALSE, out.width="100%", fig.cap="Percentage of walking trips categorized by origin and destination", fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/walking_hm_fig.jpg"))
```

For cycling trips, Figure \ref{fig:figure-02}, shows that in 1992, when
this mode of transportation was first included as an activity, the
majority of trips were from `home` to `work or school`, accounting for
about 25% of cases. This pattern remained in 2015, with these trips
representing 30% of the cases. However, a notable change occurred in
`home` to `home` trips, which decreased significantly from 19% in 1992
to 5% in 2015.

```{r  figure-02, echo=FALSE, out.width="100%", fig.cap="Percentage of cycling trips categorized by origin and destination", fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/cycling_hm_fig.jpg"))
```

# Python integration

CAN WE PROVIDE PYTHON INTEGRATION? See here:

https://github.com/paezha/idealista18/tree/master/Python

<!--

## Pre-estimated impedance functions for active travel in Canada

`ActiveCA` includes a total of
`r nrow(unique(gss_impedances[,c("YEAR","dest_label","f_name","MODE")]))`
pre-estimated impedance functions for cycling and walking trips for Census Agglomerations and Census Metropolitan Areas. Table \ref{tab:table-04} shows the best impedance functions for walking trips to `Work or school` across survey years, where all distributions, except for a gamma function in 1998, are log-normal.

```{r table-04, echo=FALSE, cache=FALSE, warning=FALSE}
# selecting mode and destination to display impedance functions
mode <- c("Walking","Cycling")
destino <- "Work or school"

# filtering the impedance functions in accordance with the mode/destination setted in the step before and creating a latex table 
selected_functions <- gss_impedances_renamed %>%
  filter(dest_label == destino, MODE %in% mode) %>%
  group_by(MODE, YEAR, dest_label) %>% 
  summarise(Function = first(f_name), 
            est_1 = first(est_1),
            est_2 = first(est_2),
            AIC = first(aic),
            f_name = first(f_name),
            count = first(count),
            .groups = "drop") %>% 
  mutate(Function = case_when(
    Function == "lnorm" ~ "Lognormal",
    Function == "gamma" ~ "Gamma",
    Function == "unif" ~ "Uniform",
    TRUE ~ Function), 
  across(c(est_1,est_2), ~ round(., digits = 2)), 
  across(c(AIC), ~ round(., digits = 0))) %>% 
  rename(Mode = MODE, Year = YEAR, Destination = dest_label) %>%
  unique() %>% 
  arrange(Mode, Year) %>%
  select(Mode, Year, Function, est_1, est_2, AIC, count)

selected_functions_tab <- selected_functions %>% 
  kable(format = "latex",
        booktabs = TRUE,
        caption = paste0("\\label{table-04}Impedance functions and AIC for `", mode ,"` trips considering `", destino,"` as destination."), 
        col.names = c("Mode", "Year", "Impedance function","Parameter 1*","Parameter 2*", "AIC","Count")) %>% 
    kable_styling(full_width = FALSE, 
                latex_options=c("scale_down"),
                font_size = 10) %>% 
  row_spec(0, bold = TRUE, align = "c") %>% 
  collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>%
  footnote(general = c("*For lognormal distributions, 'Parameter 1' and 'Parameter 2' refer to the mean and standard deviation of the distribution on the logarithmic scaler, respectively. For gamma distribution, 'Parameter 1' and 'Parameter 2' refer to the rate and shape of the distribution, respectively. `AIC` means Akaike information criterion."), threeparttable = T)

selected_functions_tab
```

Figure \ref{fig:figure-03} illustrates these fitted functions (red line)
alongside histograms of empirical travel times (grey bars). As for the
others pre-estimated functions, these examples allow to calculate gravity-based
accessibility measures for active travel across various destinations and temporal 
scales in Canadian urban areas.

```{r theoretical_impedance_values, include = FALSE}
theoretical_impedance <- function(df_functions, travel_cost) {

  # Initialize the main data frame to store results
  result <- data.frame()

  # Create the 't' sequence
  t_seq <- seq(1, travel_cost, 1)

  for (mode in unique(df_functions$Mode)){

    for (year in unique(df_functions[df_functions$Mode == mode, ]$Year)){

      for (destino in unique(df_functions[df_functions$Year == year & df_functions$Mode == mode, ]$Destination)){

      name_function <- df_functions[df_functions$Year == year & 
                                      df_functions$Mode == mode & 
                                      df_functions$Destination == destino, ]$f_name

      print(paste0("YEAR: ", year, " MODE: ", mode," Destination: ", destino))

      subset <- df_functions %>% 
        filter(Year == year & Mode == mode & Destination == destino)

      # Temporary data frame for this iteration
      temp_df <- data.frame(t = t_seq)

      # Add the appropriate distribution
      if (name_function == "lnorm") {
        temp_df$f <- dlnorm(temp_df$t,
                            meanlog = subset$est_1[1],
                            sdlog = subset$est_2[1])

      } else if (name_function == "gamma") {  
        temp_df$f <- dgamma(temp_df$t,
                            shape = subset$est_1[1],
                            rate = subset$est_2[1])

      } else if (name_function == "norm") {  
        temp_df$f <- dnorm(temp_df$t,
                           mean = subset$est_1[1],
                           sd = subset$est_2[1])

      } else if (name_function == "exp") {  
        temp_df$f <- dexp(temp_df$t,
                          rate = subset$est_1[1])

      } else if (name_function == "unif") {  
        temp_df$f <- dunif(temp_df$t,
                           min = subset$est_1[1],
                           max = subset$est_2[1])
      }

      # Add mode, year and destination columns
      temp_df$MODE <- mode
      temp_df$YEAR <- year
      temp_df$dest_label <- destino

      # Append to the result data frame
      result <- rbind(result, temp_df)
    }
  }
}
  return(result)
}

max_duration <- 100

selected_functions <- gss_impedances_renamed %>% 
  group_by(MODE, YEAR, dest_label) %>% 
  summarise(Function = first(f_name), 
            est_1 = first(est_1),
            est_2 = first(est_2),
            AIC = first(aic),
            f_name = first(f_name),
            count = first(count),
            .groups = "drop") %>% 
  mutate(Function = case_when(
    Function == "lnorm" ~ "Lognormal",
    Function == "gamma" ~ "Gamma",
    Function == "unif" ~ "Uniform",
    TRUE ~ Function), 
  across(c(est_1,est_2), ~ round(., digits = 2)), 
  across(c(AIC), ~ round(., digits = 0))) %>% 
  rename(Mode = MODE, Year = YEAR, Destination = dest_label)

theoretical_values <- theoretical_impedance(selected_functions, travel_cost = max_duration)
```

```{r include = FALSE}

for(destino in unique(gss_impedances_renamed$dest_label)){

  medianas <- gss_impedances_renamed %>%
    filter(dest_label == destino) %>%
    group_by(MODE) %>%
    summarise(mediana_duracao = median(DURATION, na.rm = TRUE), .groups = 'drop')

  maximos <- theoretical_values %>%
    filter(dest_label == destino) %>%
    group_by(MODE, YEAR) %>%
    filter(f == max(f)) %>%
    slice(which.max(t)) %>%
    ungroup()

combined_plot <- gss_impedances_renamed %>% 
  filter(dest_label == destino) %>%
  mutate(dest_label = factor(dest_label, levels = destination_order)) %>%
  ggplot(aes(x= DURATION)) +
  geom_histogram(aes(y = ..density.., colour = MODE, fill = MODE), bins = 20, alpha=0.1, position="identity") + 
  geom_line(data = theoretical_values[theoretical_values$dest_label == destino, ], aes(x = t, y = f, colour = MODE), size = 0.8) +   
  geom_vline(data = medianas, aes(xintercept = mediana_duracao, color = MODE), linetype = "dashed") + 
  geom_point(data = maximos, aes(x = t, y = f, colour = MODE), size = 3) + 
  geom_label(data = maximos, aes(x = t, y = f, label = round(t, 2), colour = MODE), 
            vjust = 0, size = 3) +
  theme_minimal() +
  facet_wrap(~YEAR) + 
  labs(x = "Trip duration (in minutes)", y = "Density", colour = "Transportation Mode", fill = "Transportation Mode") +
  theme(legend.position="right", 
    strip.text = element_text(face = "bold", hjust = 0))

  plots <- length(unique(gss_impedances_renamed[gss_impedances_renamed$dest_label == destino, ]$YEAR))

  # Case of only one year 
    if(plots == 1){
      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,".jpg"), 
       plot = combined_plot,
       width = 7, 
       height = 4, 
       units = "in", 
       dpi = 300)

    # Destination present in 3 surveys  
    } else if(plots == 3){
      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,".jpg"), 
       plot = combined_plot,
       width = 14, 
       height = 4, 
       units = "in", 
       dpi = 300)

  # Destination present in 5 surveys  
    }else {
      ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/impf_",destino,".jpg"), 
       plot = combined_plot,
       width = 14, 
       height = 9, 
       units = "in", 
       dpi = 300)
    }
}
```

```{r figure-03, echo=FALSE, out.width="100%", fig.cap=paste0("Empirical data and impedance functions fitted for `work or school` as destination."), fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/impf_Work or school.jpg"))
```

-->

# Concluding remarks

This paper presents `ActiveCA`, an open data product that provides analysis-ready 
data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) 
of the GSS surveys on active travel in Canada. In the form of an `R` data package, 
{ActiveCA} was developed after collecting, cleaning, and processing the survey data, 
providing information on origins, destinations, and duration of active travel, as
well other information.

It is important to remark that the present version of {ActiveCA} covers all Canadian
time use surveys up to 2015. While the most recent time use survey was carried out
in 2022 [@wray2024], the Public Use Microdata Files are currently unavailable, and
at the moment it is estimated that they will only be published in the later part
of 2025. The `R` package will be updated once these files become available.

<!--

Additionally, the package includes a series of pre-estimated impedance functions 
for walking and cycling trips, considering various destinations and time periods, 
for Canadian Metropolitan and Census Agglomeration areas.

-->

The value of {ActiveCA} lies in its transparency, accessibility, and ease of use,
which facilitates the addition of complementary data sets in the future. `R` users
can seamlessly explore GSS walking and cycling episodes <!-- along with calibrated impedance functions-->,
with the option to suggest enhancements to the package as needed. This article adopts
the structure proposed by Anastasia and Páez [-@soukhov2023], whose work provided
essential guidance for the creation of this package. Similarly, we aim to contribute
to the academic community by promoting transparent research practices that encourage
replication and innovation in related fields. We believe that {ActiveCA} will serve
as a basis for further research on GSS and for the integration of additional data by
the authors or the wider open source community.

# Declaration of Conflicting Interests

The author(s) declared no potential conflicts of interest with respect
to the research, authorship, and/or publication of this article.

# Funding

The author(s) disclosed receipt of the following financial support for
the research, authorship, and/or publication of this article: This work
was supported by the Social Sciences and Humanities Research Council of
Canada (*More description about the funding source after the review
process*).

# ORCID

Author 1

Author 2

Author 3

# Data availability statement

The {ActiveCA} R data package can be found and installed on Github
(*link*).

# References {#references .unnumbered}