### 2024-12-05 data-raw/03-preprocessing-main-files.Rmd
preprocessing main files (2015)

---
title: "Reading main files and preparing data sets"
author: "Bruno Santos, Mahdis Moghadasi & Antonio Paez"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: activeca.bib
link-citations: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# layout configuration 
library(tufte)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(htmltools.dir.version = FALSE)
rm(list=ls())
```

# Introduction

This Rmarkdown file is part of the
[**ActiveCA**](https://github.com/dias-bruno/ActiveCA) package. The main objective of the *ActiveCA* package is to identify the most
appropriate impedance functions for active transportation modes for
various destinations and time periods in Canada. To do achieve this objective, we analyzed historical data from the [General Social Survey
(GSS)](https://www23.statcan.gc.ca/imdb/p2SV.pl?Function=getSurvey&SDDS=5221)
from 1986 to 2015 in Canada to calculate the impedance function for
cycling and walking trips.

In this R markdown, we read the main files of GSS surveys and prepare the data set to
the following analyses.

# The General Social Surveys (GSS)

In Canada, Statistics Canada produces the General Social Surveys (GSS)
to collect data on social trends in order to track changes in the living
conditions of Canadians and well-being over time. This survey tracks
changes in time and it is used to understand how Canadians spend and
manage their time and what factors contribute to their happiness and
stress. The survey was created in 1985 as a series of independent,
annual, cross-sectional surveys, each covering one topic in depth.
Caregiving, families, time use, social identity, volunteering, and
victimization are among the current themes of the GSS.

The six survey themes listed above are repeated every five years. In
addition to the central topic, each cycle includes new content that
addresses emerging and policy-relevant issues, collecting detailed
socio-demographic information, such as age, gender, education, religion,
ethnicity, income, etc. Time-use surveys collect data on all human
activities and can inform a wide range of policies. In particular, three
key themes have been identified as essential for policymaking and for
which no other data source is adequate: unpaid work and non-market
production, well-being, and gender equality. Leisure time, work-life
balance, health, commuting, culture, and sports are among the other
topics covered by time-use surveys. In addition, statistics Canada has
been conducting time-use surveys at five to seven year intervals since
1986, with the most recent being in 2015 (1986, 1992, 1998, 2005, 2010,
and 2015). The time use collects information on participation of
respondents and time spent on a wide range of day-to-day activities
using a 24-hour retrospective diary. Also, information is collected on
the location of these activities (e.g., at home, at work, etc.) and, for
non-personal activities, the people who were present with the respondent
at the time of the activity.

The surveys (obtained from
[here](http://odesi2.scholarsportal.info/webview/)) consist of two
files: a main file and an episode file.

## The Main File

The Main File compiles an large array of aggregated data, summarizing
the answers to the questionnaire and derived variables that summarize
the respondents' time use across different activities, locations, and
social interactions. This file documents the time and duration that
respondents allocate to each activity and location. The Main File
provides a overview of daily routines and social dynamics, not focusing
on individual activity episodes. Additionally, this file categorizes
activities into bigger groups and subcategories, enhancing the data's
analytical utility with additional metrics such as total transit time,
duration spent with household members, and counts of activities and
episodes. This file is aggregated and prohibits the direct association
between specific activities with particular locations or companions.

# Reading and processing the GSS files

In this R markdown, our aim is to generate a table in the appropriate
format to perform the subsequent analyses, after reading and processing
the GSS main surveys.

To start processing the data, first load the `R` packages used in this
notebook:

```{r read-libraries}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(here) # A Simpler Way to Find Your Files
library(readxl) # Read Excel Files
library(tidyr) # Tidy Messy Data
library(usethis) # Automate Package and Project Setup
library(purrr) # Provides a complete and consistent set of tools for working with functions and vectors
```

For each survey year, the process of preparing the data is very similar
and, in the end, we will obtain a table in which each row represents an
specific respondent, and each column a variable. Since every survey introduced different variables, 
we will not try to align the name columns and values between the tables.

The file name pattern is as follows: `gss_main_` +
`year of survey` [^4]. This means that, for instance, the file
`gss_main_20105` refers to the main file of the 2005 GSS survey.

## GSS 2015

Read the Main files of the 2015 GSS survey:

```{r reading-2015-main-file}
# Main File
gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
```

Select only the socioeconomic variables:

```{r filtering-variables-2015}
gss_main_2015 <- gss_m_2015 %>%
  select(PUMFID:DVTDAY)
```

Labeling socioeconomic variables:

```{r SURVMNTH}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # SURVMNTH: Survey month of data collection from April 2015 to March 2016
    SURVMNTH = 
           case_when(SURVMNTH == 1 ~ "January", 
                     SURVMNTH == 2 ~ "February",
                     SURVMNTH == 3 ~ "March",
                     SURVMNTH == 4 ~ "April",
                     SURVMNTH == 5 ~ "May",
                     SURVMNTH == 6  ~ "June",
                     SURVMNTH == 7 ~ "July",
                     SURVMNTH == 8 ~ "August",
                     SURVMNTH == 9 ~ "September",
                     SURVMNTH == 10 ~ "October",
                     SURVMNTH == 11 ~ "November",
                     SURVMNTH == 12 ~ "Devember",
                     SURVMNTH == 96 ~ "Valid skip",
                     SURVMNTH == 97 ~ "Don't know",
                     SURVMNTH == 98 ~ "Refusal",
                     SURVMNTH == 99 ~ "Not stated"), 
    SURVMNTH = factor(SURVMNTH))

summary(gss_main_2015$SURVMNTH)
```

```{r AGEGR10}
values_ordered <- c("15 to 24 years","25 to 34 years","35 to 44 years",
                    "45 to 54 years", "55 to 64 years","65 to 74 years","75 years and over")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # AGEGR10: Age group of respondent (groups of 10)
    AGEGR10 = 
           case_when(AGEGR10 == 1 ~ "15 to 24 years", 
                     AGEGR10 == 2 ~ "25 to 34 years",
                     AGEGR10 == 3 ~ "35 to 44 years",
                     AGEGR10 == 4 ~ "45 to 54 years",
                     AGEGR10 == 5 ~ "55 to 64 years",
                     AGEGR10 == 6  ~ "65 to 74 years",
                     AGEGR10 == 7 ~ "75 years and over",
                     AGEGR10 == 96 ~ "Valid skip",
                     AGEGR10 == 97 ~ "Don't know",
                     AGEGR10 == 98 ~ "Refusal",
                     AGEGR10 == 99 ~ "Not stated"), 
    AGEGR10 = factor(AGEGR10, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$AGEGR10)
```

```{r SEX}
values_ordered <- c("Male", "Female")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # SEX: Sex of respondent
    SEX = 
           case_when(SEX == 1 ~ "Male", 
                     SEX == 2 ~ "Female",
                     SEX == 96 ~ "Valid skip",
                     SEX == 97 ~ "Don't know",
                     SEX == 98 ~ "Refusal",
                     SEX == 99 ~ "Not stated"), 
    SEX = factor(SEX, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$SEX)
```

```{r MARSTAT}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # MARSTAT: Marital status of the respondent
    MARSTAT = 
           case_when(MARSTAT == 1 ~ "Married", 
                     MARSTAT == 2 ~ "Living common-law",
                     MARSTAT == 3 ~ "Widowed",
                     MARSTAT == 4 ~ "Separated",
                     MARSTAT == 5 ~ "Divorced",
                     MARSTAT == 6  ~ "Single, never married",
                     MARSTAT == 96 ~ "Valid skip",
                     MARSTAT == 97 ~ "Don't know",
                     MARSTAT == 98 ~ "Refusal",
                     MARSTAT == 99 ~ "Not stated"), 
    MARSTAT = factor(MARSTAT))

summary(gss_main_2015$MARSTAT)
```

```{r AGEPRGRD}
values_ordered <- c("Respondent is 11 + years older", 
                    "Respondent is 6 - 10 years older",
                    "Respondent is 5 years older",
                    "Respondent is 4 years older",
                    "Respondent is 3 years older",
                    "Respondent is 2 years older",
                    "Respondent is 1 years older",
                    "Respondent and spouse/partner are the same age",
                    "Respondent is 1 years younger",
                    "Respondent is 2 years younger",
                    "Respondent is 3 years younger",
                    "Respondent is 4 years younger",
                    "Respondent is 5 years younger",
                    "Respondent is 6 - 10 years younger",
                    "Respondent is 11 + years younger",
                    "Valid skip")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # AGEPRGRD: Age difference between respondent and spouse/partner.
    AGEPRGRD = 
           case_when(AGEPRGRD == 1 ~ "Respondent is 11 + years older", 
                     AGEPRGRD == 2 ~ "Respondent is 6 - 10 years older",
                     AGEPRGRD == 3 ~ "Respondent is 5 years older",
                     AGEPRGRD == 4 ~ "Respondent is 4 years older",
                     AGEPRGRD == 5 ~ "Respondent is 3 years older",
                     AGEPRGRD == 6  ~ "Respondent is 2 years older",
                     AGEPRGRD == 7  ~ "Respondent is 1 years older",
                     AGEPRGRD == 8  ~ "Respondent and spouse/partner are the same age",
                     AGEPRGRD == 9  ~ "Respondent is 1 years younger",
                     AGEPRGRD == 10  ~ "Respondent is 2 years younger",
                     AGEPRGRD == 11  ~ "Respondent is 3 years younger",
                     AGEPRGRD == 12  ~ "Respondent is 4 years younger",
                     AGEPRGRD == 13  ~ "Respondent is 5 years younger",
                     AGEPRGRD == 14  ~ "Respondent is 6 - 10 years younger",
                     AGEPRGRD == 15  ~ "Respondent is 11 + years younger",
                     AGEPRGRD == 96 ~ "Valid skip",
                     AGEPRGRD == 97 ~ "Don't know",
                     AGEPRGRD == 98 ~ "Refusal",
                     AGEPRGRD == 99 ~ "Not stated"), 
    AGEPRGRD = factor(AGEPRGRD, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2015$AGEPRGRD)
```

```{r SEXPR}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # SEXPR: Sex of respondent's spouse/partner living in the household.
    SEXPR = 
           case_when(SEXPR == 1 ~ "Male spouse/partner in the household", 
                     SEXPR == 2 ~ "Female spouse/partner in the household",
                     SEXPR == 6 ~ "Valid skip",
                     SEXPR == 7 ~ "Don't know",
                     SEXPR == 8 ~ "Refusal",
                     SEXPR == 9 ~ "Not stated"), 
    SEXPR = factor(SEXPR))

summary(gss_main_2015$SEXPR)
```

```{r PRTYPEC}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # PRTYPEC: Type of partner respondent has living in the household.
    PRTYPEC = 
           case_when(PRTYPEC == 1 ~ "Opposite sex married partner in the household", 
                     PRTYPEC == 2 ~ "Opposite sex common-law partner in the household",
                     PRTYPEC == 3 ~ "Same sex married or common-law partner in the household",
                     PRTYPEC == 6 ~ "Valid skip",
                     PRTYPEC == 7 ~ "Don't know",
                     PRTYPEC == 8 ~ "Refusal",
                     PRTYPEC == 9 ~ "Not stated"), 
    PRTYPEC = factor(PRTYPEC))

summary(gss_main_2015$PRTYPEC)
```

```{r CXRFLAG}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # CXRFLAG:  Child(ren) of the respondent living in the household.
    CXRFLAG = 
           case_when(CXRFLAG == 1 ~ "Yes", 
                     CXRFLAG == 2 ~ "No",
                     CXRFLAG == 6 ~ "Valid skip",
                     CXRFLAG == 7 ~ "Don't know",
                     CXRFLAG == 8 ~ "Refusal",
                     CXRFLAG == 9 ~ "Not stated"), 
    CXRFLAG = factor(CXRFLAG))

summary(gss_main_2015$CXRFLAG)
```

```{r AGECXRYG}
values_ordered <- c( "Valid skip", "5 to 4", "5 to 9","10 to 14",
                     "15 to 19","20 to 24",
                     "25 years and older")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # AGECXRYG: Age of respondent's youngest child in household
    AGECXRYG = 
           case_when(AGECXRYG == 1 ~ "5 to 4", 
                     AGECXRYG == 2 ~ "5 to 9",
                     AGECXRYG == 3 ~ "10 to 14",
                     AGECXRYG == 4 ~ "15 to 19",
                     AGECXRYG == 5 ~ "20 to 24",
                     AGECXRYG == 6  ~ "25 years and older",
                     AGECXRYG == 96 ~ "Valid skip",
                     AGECXRYG == 97 ~ "Don't know",
                     AGECXRYG == 98 ~ "Refusal",
                     AGECXRYG == 99 ~ "Not stated"), 
    AGECXRYG = factor(AGECXRYG, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$AGECXRYG)
```
```{r CHINHSDC}
values_ordered <- c("None", 
                     "One",
                     "Two",
                     "Three",
                     "Four or more")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # CHINHSDC: Respondent’s child(ren) in household - Any age/marital status
    CHINHSDC = 
           case_when(CHINHSDC == 0 ~ "None", 
                     CHINHSDC == 1 ~ "One",
                     CHINHSDC == 2 ~ "Two",
                     CHINHSDC == 3 ~ "Three",
                     CHINHSDC == 4 ~ "Four or more",
                     CHINHSDC == 6 ~ "Valid skip",
                     CHINHSDC == 7 ~ "Don't know",
                     CHINHSDC == 8 ~ "Refusal",
                     CHINHSDC == 9 ~ "Not stated"), 
    CHINHSDC = factor(CHINHSDC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$CHINHSDC)
```

```{r CXR0014C}
values_ordered <- c("None", 
                     "One",
                     "Two",
                     "Three or more")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # CXR0014C:  Respondent's child(ren) in household - 0 to 14 years
    CXR0014C = 
           case_when(CXR0014C == 0 ~ "None", 
                     CXR0014C == 1 ~ "One",
                     CXR0014C == 2 ~ "Two",
                     CXR0014C == 3 ~ "Three or more",
                     CXR0014C == 6 ~ "Valid skip",
                     CXR0014C == 7 ~ "Don't know",
                     CXR0014C == 8 ~ "Refusal",
                     CXR0014C == 9 ~ "Not stated"), 
    CXR0014C = factor(CXR0014C, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$CXR0014C)
```

```{r CHH0014C}
values_ordered <- c("None", 
                     "One",
                     "Two",
                     "Three or more")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # CHH0014C:  Child(ren) in household - 0 to 14 years
    CHH0014C = 
           case_when(CHH0014C == 0 ~ "None", 
                     CHH0014C == 1 ~ "One",
                     CHH0014C == 2 ~ "Two",
                     CHH0014C == 3 ~ "Three or more",
                     CHH0014C == 6 ~ "Valid skip",
                     CHH0014C == 7 ~ "Don't know",
                     CHH0014C == 8 ~ "Refusal",
                     CHH0014C == 9 ~ "Not stated"), 
    CHH0014C = factor(CHH0014C, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$CHH0014C)
```

```{r CXRTIME6}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # CXRTIME6:  Age group of respondent’s child(ren) in household
    CXRTIME6 = 
           case_when(CXRTIME6 == 1 ~ "No child under 19 years of age at home",
                     CXRTIME6 == 2 ~ "All children under 5 years of age",
                     CXRTIME6 == 3 ~ "All children between 5 and 12 years of age",
                     CXRTIME6 == 4 ~ "All children 13 years of age and older",
                     CXRTIME6 == 5 ~ "At least one child under 5 years of age but not all
children",
                     CXRTIME6 == 6 ~ "Other",
                     CXRTIME6 == 96 ~ "Valid skip",
                     CXRTIME6 == 97 ~ "Don't know",
                     CXRTIME6 == 98 ~ "Refusal",
                     CXRTIME6 == 99 ~ "Not stated"), 
    CXRTIME6 = factor(CXRTIME6))

summary(gss_main_2015$CXRTIME6)
```

```{r PARNUM}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # PARNUM: Number of parents the respondent has in household
    PARNUM = 
           case_when(PARNUM == 0 ~ "None",
                     PARNUM == 1 ~ "One parent",
                     PARNUM == 2 ~ "Two parents",
                     PARNUM == 6 ~ "Valid skip",
                     PARNUM == 7 ~ "Don't know",
                     PARNUM == 8 ~ "Refusal",
                     PARNUM == 9 ~ "Not stated"), 
    PARNUM = factor(PARNUM))

summary(gss_main_2015$PARNUM)
```

```{r LIVARR08}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # LIVARR08: Living arrangement of respondent’s household (8 categories)
    LIVARR08 = 
           case_when(LIVARR08 == 1 ~ "Under 25 years - Not living with parents",
                     LIVARR08 == 2 ~ "Under 25 years - Living with both parents",
                     LIVARR08 == 3 ~ "Under 25 years - Living with mother only",
                     LIVARR08 == 4 ~ "Under 25 years - Living with father only",
                     LIVARR08 == 5 ~ "25 years of age or older - Not living with parents",
                     LIVARR08 == 6 ~ "25 years of age or older - Living with both parents",
                     LIVARR08 == 7 ~ "25 years of age or older - Living with mother only",
                     LIVARR08 == 8 ~ "25 years of age or older - Living with father only",
                     LIVARR08 == 96 ~ "Valid skip",
                     LIVARR08 == 97 ~ "Don't know",
                     LIVARR08 == 98 ~ "Refusal",
                     LIVARR08 == 99 ~ "Not stated"), 
    LIVARR08 = factor(LIVARR08))

summary(gss_main_2015$LIVARR08)
```

```{r LIVARR11} 
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # LIVARR11: Living arrangement of respondent’s household (11 categories)
    LIVARR11 = 
           case_when(LIVARR11 == 1 ~ "Living alone",
                     LIVARR11 == 2 ~ "Living with sponse only",
                     LIVARR11 == 3 ~ "Living with spouse and child(ren) < 25 years",
                     LIVARR11 == 4 ~ "Living with spouse and child(ren) 25 years or older",
                     LIVARR11 == 5 ~ "Living with spouse and other",
                     LIVARR11 == 6 ~ "Living without spouse but with chil(ren) < 25 years",
                     LIVARR11 == 7 ~ "Living without spouse but with chil(ren) 25 years and older",
                     LIVARR11 == 8 ~ "Living with two parents",
                     LIVARR11 == 9 ~ "Living with one parent",
                     LIVARR11 == 10 ~ "Multiple person household - Other living arrangement",
                     LIVARR11 == 11 ~ "Multiple person household - Relationships undefined",
                     LIVARR11 == 96 ~ "Valid skip",
                     LIVARR11 == 97 ~ "Don't know",
                     LIVARR11 == 98 ~ "Refusal",
                     LIVARR11 == 99 ~ "Not stated"), 
    LIVARR11 = factor(LIVARR11))

summary(gss_main_2015$LIVARR11)
```

```{r HSDSIZEC}
values_ordered <- c("One person household",
                   "Two person household",
                   "Three person household",
                   "Four person household",
                   "Five person household",
                   "Six or more person household")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # HSDSIZEC: Household size of respondent
    HSDSIZEC = 
           case_when(HSDSIZEC == 1 ~ "One person household",
                     HSDSIZEC == 2 ~ "Two person household",
                     HSDSIZEC == 3 ~ "Three person household",
                     HSDSIZEC == 4 ~ "Four person household",
                     HSDSIZEC == 5 ~ "Five person household",
                     HSDSIZEC == 6 ~ "Six or more person household",
                     HSDSIZEC == 96 ~ "Valid skip",
                     HSDSIZEC == 97 ~ "Don't know",
                     HSDSIZEC == 98 ~ "Refusal",
                     HSDSIZEC == 99 ~ "Not stated"), 
    HSDSIZEC = factor(HSDSIZEC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$HSDSIZEC)
```

```{r HSDELIGC}
values_ordered <- c("One member",
                     "Two members",
                     "Three members",
                     "Four or more members")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # HSDELIGC: Number of members in respondent’s household 15 years of age or older
    HSDELIGC = 
           case_when(HSDELIGC == 1 ~ "One member",
                     HSDELIGC == 2 ~ "Two members",
                     HSDELIGC == 3 ~ "Three members",
                     HSDELIGC == 4 ~ "Four or more members",
                     HSDELIGC == 6 ~ "Valid skip",
                     HSDELIGC == 7 ~ "Don't know",
                     HSDELIGC == 8 ~ "Refusal",
                     HSDELIGC == 9 ~ "Not stated"), 
    HSDELIGC = factor(HSDELIGC, levels = values_ordered, ordered = TRUE))

summary(gss_main_2015$HSDELIGC)
```

```{r MULTIGEN}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # MULTIGEN:  Three generations or more in the respondent’s household
    MULTIGEN = 
           case_when(MULTIGEN == 1 ~ "Yes", 
                     MULTIGEN == 2 ~ "No",
                     MULTIGEN == 6 ~ "Valid skip",
                     MULTIGEN == 7 ~ "Don't know",
                     MULTIGEN == 8 ~ "Refusal",
                     MULTIGEN == 9 ~ "Not stated"), 
    MULTIGEN = factor(MULTIGEN))

summary(gss_main_2015$MULTIGEN)
```

```{r PRV}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # PRV: Province of residence
    PRV = 
      case_when(PRV == 10 ~ "Newfoundland and Labrador",
                     PRV == 11 ~ "Prince Edward Island",
                     PRV == 12 ~ "Nova Scotia",
                     PRV == 13 ~ "New Brunswick",
                     PRV == 24 ~ "Quebec",
                     PRV == 35 ~ "Ontario",
                     PRV == 46 ~ "Manitoba",
                     PRV == 47 ~ "Saskatchewan",
                     PRV == 48 ~ "Alberta",
                     PRV == 59 ~ "British Columbia",
                     PRV == 96 ~ "Valid skip",
                     PRV == 97 ~ "Don't know",
                     PRV == 98 ~ "Refusal",
                     PRV == 99 ~ "Not stated"),
    PRV = factor(PRV))

summary(gss_main_2015$PRV) 
```

```{r LUC_RST}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # LUC_RST:  Population centre indicator
    LUC_RST = 
      case_when(LUC_RST == 1 ~ "CMA/CA",
                     LUC_RST == 2 ~ "non CMA/CA",
                     LUC_RST == 3 ~ "Prince Edward Island",
                     LUC_RST == 6 ~ "Valid skip",
                     LUC_RST == 7 ~ "Don't know",
                     LUC_RST == 8 ~ "Refusal",
                     LUC_RST == 9 ~ "Not stated"),
    LUC_RST = factor(LUC_RST))

summary(gss_main_2015$LUC_RST)
```

```{r GTU_110}
values_ordered <- c("Every day",
                    "A few times a week",
                    "About once a week",
                    "About once a month",
                    "Less than once a month",
                    "Never", 
                    "Don't know")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # GTU_110: How often do you feel rushed? Would you say it is...?
    GTU_110 = 
      case_when(GTU_110 == 1 ~ "Every day",
                     GTU_110 == 2 ~ "A few times a week",
                     GTU_110 == 3 ~ "About once a week",
                     GTU_110 == 4 ~ "About once a month",
                     GTU_110 == 5 ~ "Less than once a month",
                     GTU_110 == 6 ~ "Never",
                     GTU_110 == 96 ~ "Valid skip",
                     GTU_110 == 97 ~ "Don't know",
                     GTU_110 == 98 ~ "Refusal",
                     GTU_110 == 99 ~ "Not stated"),
    GTU_110 = factor(GTU_110, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2015$GTU_110)
```

```{r GTU_130}
values_ordered <- c("Every day",
                    "A few times a week",
                    "About once a week",
                    "About once a month",
                    "Less than once a month",
                    "Never", 
                    "Don't know")

gss_main_2015 <- gss_main_2015 %>% 
  mutate( # GTU_130: How often do you feel you have time on your hands that you don’t know what to do with?
    GTU_130 = 
      case_when(GTU_130 == 1 ~ "Every day",
                     GTU_130 == 2 ~ "A few times a week",
                     GTU_130 == 3 ~ "About once a week",
                     GTU_130 == 4 ~ "About once a month",
                     GTU_130 == 5 ~ "Less than once a month",
                     GTU_130 == 6 ~ "Never",
                     GTU_130 == 96 ~ "Valid skip",
                     GTU_130 == 97 ~ "Don't know",
                     GTU_130 == 98 ~ "Refusal",
                     GTU_130 == 99 ~ "Not stated"),
    GTU_130 = factor(GTU_130, levels = rev(values_ordered), ordered = TRUE))

summary(gss_main_2015$GTU_130)
```

```{r DVTDAY}
gss_main_2015 <- gss_main_2015 %>% 
  mutate( # DVTDAY: Diary - Reference day - Type
    DVTDAY = 
      case_when(DVTDAY == 1 ~ "Weekday",
                     DVTDAY == 2 ~ "Saturday",
                     DVTDAY == 3 ~ "Sunday",
                     DVTDAY == 96 ~ "Valid skip",
                     DVTDAY == 97 ~ "Don't know",
                     DVTDAY == 98 ~ "Refusal",
                     DVTDAY == 99 ~ "Not stated"),
    DVTDAY = factor(DVTDAY))

summary(gss_main_2015$DVTDAY)
```
Save data in R format for later use:
```{r}
usethis::use_data(cycling_2015, 
                  overwrite = TRUE)
```

## GSS 2010

Read the Main files of the 2010 GSS survey:

```{r reading-2010-main-file}
# Main File
gss_m_2010 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2010/gss-12M0018-E-2010-c-24-tus-mf_F1.csv"))
```

Select only the socioeconomic variables:

```{r filtering-variables-2010}
gss_main_2010 <- gss_m_2010 %>%
  select(RECID:DVTDAY)
```