### 2025-05-31 paper/Manuscript-Data-Package-1Round-Review.Rmd
Paper revision.

---
title: "ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel"
author:
  - name: "Anon1"
    email: "Anon1@anon.org"
    affiliation: CODE
    footnote: 1
  - name: "Anon2"
    email: "Anon2@anon.org"
    affiliation: CODE
  - name: "Anon3"
    email: "Anon3@anon.org"
    affiliation: CODE
address:
  - code: CODE
    address: "Some school"
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
  "This paper describes {ActiveCA}, an open data product with Canadian time use data. {ActiveCA} is an `R` data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) cicles from the General Social Survey (GSS). Active travel is characterized by mode, with walking being part of every cycle and bicycling starting in 1992. The attributes of active trips are the types of locations of origins and destinations, the duration of trips, and episode weights for expanding the trips to population-wide estimates. Based on the year of the survey, a variety of locations are coded. In earlier cycles, these include home, work or school, and other's home, whereas in later cycles these are augmented with locations such as grocery stores, restaurants, outdoor destinations, and others. The geographical resolution includes the province and whether the episode was in an urban or rural setting."

keywords: Active; mobility; walking; cycling; travel time; time-use; General Social Survey 
classoption:
  - Royal
  - times
bibliography: bibfile.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  warning = FALSE,
  comment = '', 
  out.width = "1\\linewidth")

knitr::knit_hooks$set(inline = function(x) {
  if (is.numeric(x)) {
    format(x, big.mark = ",", scientific = FALSE)
  } else {
    x
  }
})

opts <- options(knitr.kable.NA = "")
```

```{r install-data-package, include=FALSE}
if (!require("ActiveCA", character.only = TRUE)) {
      remotes::install_github("dias-bruno/ActiveCA",
                        build_vignettes = TRUE)
}

rm(list=ls())
```

```{r load-packages, include=FALSE, cache=FALSE}
# Load packages
library(ActiveCA)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(reshape2)
library(rlang)
library(scales)
library(gridExtra)
library(ggpubr)
library(grid)
library(stats)
library(Hmisc)
```

```{r load-data, include = FALSE}
# Load data
data("gss_episodes")
data("gss_impedances")

# Original main and _Episode files_ from 2015 to display table examples
gss_m_2015 <- read.csv(paste0(here(),
                              "/data-raw/source-files/Time-use-2015/gss-m.csv"))
gss_e_2015 <- read.csv(paste0(here(), 
                              "/data-raw/source-files/Time-use-2015/gss-e.csv"))

# Processed _Main files_
data("gss_main_2022")
data("gss_main_2015")
data("gss_main_2010")
data("gss_main_2005")
data("gss_main_1998")
data("gss_main_1992")
data("gss_main_1986")
```

```{r rename-labels, include = FALSE}
# Renaming 'dest_label' and 'orig_label' values to reduce table space
gss_episodes_renamed <- gss_episodes |>
  mutate(dest_label = case_when(
    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
    dest_label == "Restaurant, bar or club" ~ "Restaurant",
    dest_label == "Sports centre, field or arena" ~ "Sport area",
    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
    dest_label == "In the neighbourhood" ~ "Neighbourhood",
    dest_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ dest_label  # Keep all other values as they are
  ), 
  orig_label = case_when(
    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
    orig_label == "Restaurant, bar or club" ~ "Restaurant",
    orig_label == "Sports centre, field or arena" ~ "Sport area",
    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
    orig_label == "In the neighbourhood" ~ "Neighbourhood",
    orig_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ orig_label))

gss_impedances_renamed <- gss_impedances |>
  mutate(dest_label = case_when(
    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
    dest_label == "Restaurant, bar or club" ~ "Restaurant",
    dest_label == "Sports centre, field or arena" ~ "Sport area",
    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
    dest_label == "In the neighbourhood" ~ "Neighbourhood",
    dest_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ dest_label  # Keep all other values as they are
  ), 
  orig_label = case_when(
    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
    orig_label == "Restaurant, bar or club" ~ "Restaurant",
    orig_label == "Sports centre, field or arena" ~ "Sport area",
    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
    orig_label == "In the neighbourhood" ~ "Neighbourhood",
    orig_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ orig_label))
```

# Introduction

The objective of this paper is to introduce {ActiveCA}, an open data product with data from all Time Use Survey (TUS) cycles of the General Social Surveys (GSSs). Open data products
(ODPs) are the outcome of a process that transforms raw data (open or not) into 
analysis-ready data, following a transparent process in which all stages of
development follow open principles [@arribas-bel2021]. ODPs, while still open, 
differ from general open data in their degree of ease of access, their heightened 
usability, and potentially the value they add to the raw data.

{ActiveCA} provides analysis-ready data concerning active travel in Canada spanning
a period of almost 40 years, obtained from the TUS cycles of the Canadian GSS.
The GSS program is designed to provide cross-sectional data on topics of interest to improve the well-being of Canadians. As part of this 
program, every five to seven years the survey is done on the topic of time use (TUS). 
Concretely, {ActiveCA} covers Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005),
24 (2010), 29 (2015), and 34 (2022) of the TUS. Time use data in these surveys is coded using a very fine grain, from time spent in chores, leisure, and sleeping, to time spent
working or at school. These surveys have proved valuable in investigations of 
mobility and quality of life [@spinneyTransport2009], the relationship between
active travel and transit use [@lachapelleLonger2016], and travel behavior and time
poverty [@kimFacing2024], to name but a few examples.

Using Public Use Microdata Files (PUMFs) from the Time Use Survey (TUS), we extracted all data necessary to characterize active travel in Canada - specifically, episodes in which the activity involved moving between an origin and a destination by walking or cycling. Although Statistics Canada provides PUMFs and accompanying documentation for the GSS program [see @statisticscanada2024], accessing and preparing these files for analysis is not a straightforward task due to their size and complexity.  The  process 
of extracting information of interest from the source files is  time-consuming, 
tedious, and challenging and/or prone to error due to the expertise required to 
work with these files. To create {ActiveCA} we selected, labelled, and processed
the TUS cycles to make them ready for analysis.

{ActiveCA} is distributed as an `R` package with a number of data 
objects and their documentation. `R` packages contain code, data, and documentation
in a standardized format that can be installed by `R` users via a software repository,
such as CRAN (Comprehensive R Archive Network) or GitHub, which makes them an 
adroit medium to distribute analysis-ready data.

Given the level of interest in active travel [e.g., @mccurdySupport2023], reducing
the barriers to using data contained in rich, but difficult to access and use surveys,
such as TUS, is a worthy endeavour that can only improve data-driven 
decisions in transportation, urban, and health policy.

The rest of this paper discusses the sources of data, and the process implemented
to retrieve and package them. Then, we explain how to use the package and show
some selected examples of analysis to whet the imagination of potential users.
This ODP provides not only data that are easy to use, but also all the code and 
documentation that make this a reproducible research project. In summary, {ActiveCA}
aims to implement and inspire the best principles of open spatial sciences 
[@paez_open_2021; @brunsdon_opening_2021].

# The Time Use Survey (TUS) collection

Statistics Canada [-@statisticscanada2024] conducts GSS surveys to obtain data 
on social trends to track changes in Canadians' living conditions and well-being
over time. The series of survey on time use are used to understand how Canadian
residents spend and manage their time, and what factors contribute to their 
happiness and stress. The GSS program was created in 1985, and is serialized to 
provide a collection of annual, representative cross-sectional surveys.

The topics of the survey cycle every few years to cover topics that include family,
health, social identity, and every five to seven years time use. The first Canadian
TUS done as part of the GSS program was conducted in 1986, and the 
most recent was completed in 2022. These Time Use Surveys [@statisticscanada2022]
collect data on respondents' participation and time spent on a wide range of 
everyday activities using a 24-hour retrospective diary, with information on the 
location of these activities (e.g. at home, at work, etc.) and, for non-personal
activities, the people who were present with the respondent at the time of the 
activity. In addition, time-use surveys also cover topics related to leisure time,
work-life balance, health, commuting, culture and sports, and many others.

TUS allows researchers to identify the origin and destination of trips, travel times and modes of transport used, providing a valuable dataset for analyzing active travel behavior. It also provides the empirical basis for tools used in transportation analysis, such as the development of impedance functions for accessibility analysis, a measure of the ease with which people can reach destinations and opportunities [@hansen1959]. The Canadian TUS is unique at the national level in collecting detailed information on travel behaviour. Its consistent application across survey cycles enables the identification of long-term trends, with some questions present in the questionnaires since the first application of the survey.

Most respondents to the 2022 TUS completed it online, reflecting Statistics Canada's effort to adapt to technological changes and growing time demands by offering greater flexibility and convenience [@statisticscanada2022]. While such methodological changes may affect data comparability over time, it is not possible to determine whether observed differences result from actual population changes or shifts in data collection methods. Despite rigorous efforts to ensure data quality, the use of electronic questionnaires may have influenced estimates. Statistics Canada assessed the impact of collection mode on a limited set of key questions, constrained by sample size. Importantly, none of the variables used in this research are `R` package in the 2022 PUMF User Guide as unsuitable for trend analysis.

Until 2022, Statistics Canada employed a telephone-based sampling frame, which was replaced by a dwelling-based frame in the most recent cycle. Each survey cycle spans a 12-month period, typically from July to the following July. The target population includes all Canadians aged 15 and over, excluding residents of the Yukon, Northwest Territories, and Nunavut, full-time institutional residents, and individuals living on Indigenous reserves.

The survey encompasses both rural and urban areas, including metropolitan and non-metropolitan regions, to ensure a diverse and representative sample. For sampling, the ten provinces were divided into geographic strata. Several Census Metropolitan Areas (CMAs) - such as St. John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa, Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and Vancouver - were treated as separate strata. Additional strata grouped other CMAs within Quebec, Ontario, and British Columbia, as well as non-CMA areas within each province.

The Public Use Microdata are released by Statistics Canada in two files: a 
_Main file_ and an _Episode file_. The files are linked by keys that identify 
households, individuals, and episodes (i.e., activities) conducted by individuals.
We discuss these files in more detail in the following section.

## The _Main file_

The _Main file_ of the time use survey compiles a large array of aggregated data, 
summarizing the answers to the questionnaire that describe households and 
individuals, as well as derived variables that summarize the respondents' use of 
time use across different activities, locations, and social interactions. This 
file documents the time and duration that respondents allocate to each activity 
and location. The _Main file_ provides a overview of daily routines and social 
dynamics, not focusing on individual activity episodes. Additionally, this file 
categorizes activities into bigger groups and subcategories, facilitating the 
data's analytical utility with additional metrics such as total transit time, time 
spent with household members, and counts of activities and episodes.

Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the TUS PUMF 2015 _Main file_ (Cycle 29). Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (`PUMFID`), the person's weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`), the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and the respondent's marital status (`MARSTAT`).

The _Main file_ of the 2015 GSS surveys includes a total of 17,390 respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and 848 variables.  For discrete variables, Statistics Canada has assigned specific codes to the possible values, with each code accompanied by a label. For instance, in the case of the variable `SURVMNTH`, a value of `1` means `January 2016`, 2 means `February 2016`, 3 corresponds to `March 2016`, and so on.

As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not labeled. Additionally, the format of the tables (comma-separated values) does not allow for the specification of variable types (whether a variable is continuous or discrete), which can lead to mistakes analysts who have limited experience working with PUMFs.

```{r gss-main-file-2015}
main_2015_unprocessed <- gss_m_2015[1:10, 1:6]  |>
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main file of the 2015 GSS.") |>
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) |>
  footnote(general = c("Legend: `PUMFID`: record identification. `WGHT_PER`:  person weight. `SURVMNTH`: survey month of data collection. `AGEGR10`: age group of the respondent. `SEX`: sex of the respondent. `MARSTAT`: marital status of the respondent."), threeparttable = T)

main_2015_unprocessed
```

## The _Episode file_

The episode is a much bigger file that records detailed data for each activity 
episode reported by respondents. Each episode represents a single activity and its
duration, and the sum of all episodes throughout the day adds up to 24 hours. Each
entry in this file includes the start and end times of the activity, the duration,
location, and accompanying social context, informing when and where activities 
occurred and with whom. The focus of the _Episode file_ is not on the characteristics
of the respondents, but on the characteristics of the activities, and the data 
are structured around the numerous activity instances that compose a day of the 
respondent. Although respondent-specific characteristics are not included within
the _Episode file_, it is possible to link the _Main file_ and the _Episode file_ by 
using a key present in both the _Main_ and _Episode files_.

Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the _Main file_ structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the TUS PUMF 2015 _Episode file_ (Cycle 29). Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), episode number (`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and episode location (`LOCATION`).

In total, the _Episode file_ of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables detailing the episodes. Similar to the _Main file_, Statistics Canada has created codes for the discrete variables, with each value corresponding to a label.

In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this respondent began the diary description by sleeping at home (`TUI_01 = 1` and `LOCATION = 300`) for 210 minutes, followed by 40 minutes of personal hygiene (`TUI_01 = 2`). The respondent then spent 15 minutes on personal care activities, such as getting ready for school, supervising homework, reading, playing, reprimanding, or providing educational or emotional support, as indicated by `TUI_01 = 27`. Next, they recorded a travel episode, walking for 15 minutes (`TUI_01 = 7` and `LOCATION = 315`), where both the origin and destination were their home. Such trips, where the journey starts and finishes at home, can be classified as recreational or leisure trips. Next, the respondent spent 3 hours searching for a job (`TUI_01 = 9`), took a 15-minute lunch break (`TUI_01 = 6`), and then cleaned the house (`TUI_01 = 18`) for two hours. Table \ref{tab:ep-2015-unprocessed} displays only six variables out of the 527 available. As shown, since the dataset does not label the variable values, decoding them can be both time-consuming and challenging.

```{r gss-epi-file-2015}
ep_2015_unprocessed <- gss_e_2015 |>
  filter(PUMFID == 10041 & EPINO <= 7) |>
  dplyr::select(PUMFID, EPINO,WGHT_EPI,TUI_01, DURATION, LOCATION) |>
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.") |>
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) |>
  footnote(general = c("Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT_EPI: episode's weight. TUI_01: activity code. DURATION: episode's duration. LOCATION: episode's location."), threeparttable = T)

ep_2015_unprocessed
```

# Data process

Figure \ref{fig:process-figure} presents a diagram illustrating the processes applied to the _Main_ and _Episode files_ to create the {ActiveCA} datasets. For each TUS cycle, we reviewed the _Episode files_ to identify movement episodes involving walking or cycling. This allowed us to also select the activities immediately before and after the movement episode, helping to infer the trip’s purpose and determine its origin and destination. Active trips were identified by their corresponding activity codes, accounting for variations across survey cycles.

Next, we labeled the coded variables with their appropriate descriptions, classifying each origin and destination, mode of travel, and time spent in the active trip. Categorical variables were converted into factor variables, and ordinal variables were defined as ordered factors.

For the _Main files_, we selected socioeconomic variables to help profile individuals engaged in active travel episodes. These included key indicators such as age group, sex, marital status, and number of children, among others. As with the _Episode files_, we labeled and factored the coded variables. Using the appropriate identifiers, we joined the Episode and _Main files_ to determine the province and the type of population centre - whether it is a CMA, a Census Agglomeration (CA), or a non-CMA/CA area.

As output, the {ActiveCA} package provides processed datasets of walking and cycling episodes by year, a unified dataset that contains all walking and cycling episodes across TUS cycles, and processed _Main files_ (not unified, as socioeconomic and demographic variables and their categories differ across TUS cycles).

```{r process-figure, echo=FALSE, out.width="100%", fig.cap="Diagram with the processes applied to the main (blue arrows) and episode files (green arrows) to obtain the ActiveCA datasets.", fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/RPackages - ActiveCA.jpeg"))
```

# {ActiveCA} data sets

This section presents some potential applications of the {ActiveCA} `R` package. In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses. The installation instruction and also some examples of application of the {ActiveCA} `R` package are available in the `vignettes`, available in the Github repository.

## Active episodes

```{r processed-files, include=FALSE}
df_main_processed <- data.frame(
  'Survey' = c("Main", "Main","Main","Main","Main","Main","Main"),
  'Year' = c(2022, 2015, 2010, 2005, 1998, 1992, 1986),
  'Count' = c(nrow(gss_main_2022),
              nrow(gss_main_2015),
              nrow(gss_main_2010),
              nrow(gss_main_2005),
              nrow(gss_main_1998),
              nrow(gss_main_1992),
              nrow(gss_main_1986)),
  'Weighted' = c(sum(gss_main_2022$WGHT_PER),
              sum(gss_main_2015$WGHT_PER),
              sum(gss_main_2010$WGHT_PER),
              sum(gss_main_2005$WGHT_PER),
              sum(gss_main_1998$WGHTFIN),
              sum(gss_main_1992$FWGHT),
              sum(gss_main_1986$FWGT_OS/10000)))

df_ep_processed <- gss_episodes |> 
  group_by(YEAR) |>
  summarise(Survey = 'Episode', 
            Count = n(), 
            Weighted = sum(WGHT_EPI)) |>
  rename(Year = YEAR) |>
  dplyr::select(Survey, Year, Count, Weighted) |> 
  arrange(-Year)

df_processed <- rbind(df_main_processed,df_ep_processed)

# For table

df_ep_processed_table <- gss_episodes |> 
  group_by(YEAR) |>
  summarise(Survey_ep = 'Episode', 
            Count_ep = n(), 
            Weighted_ep = sum(WGHT_EPI)) |>
  rename(Year_ep = YEAR) |>
  dplyr::select(Survey_ep, Year_ep, Count_ep, Weighted_ep) |> 
  arrange(-Year_ep)

df_processed_table <- cbind(df_main_processed,df_ep_processed_table)
```

Table \ref{tab:processed-obs} displays the total number of records processed for Main and _Episode files_. For the _Main files_, a total of `r sum(df_main_processed$Count)` records were processed, referring to all records from the Time Use Surveys from 1986 to 2022, that together represents more of `r sum(df_main_processed$Weighted)` respondents. It also presents the total cases of active trips episodes identified. In total `r sum(df_ep_processed$Count)` records with register of active travel activity. Together, these records account for `r sum(df_ep_processed$Weighted)` episodes.

```{r table_df_processed}

table_df_processed <- df_processed_table |> 
  select(-c(Survey_ep, Survey, Year_ep)) %>% 
  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
        caption = "\\label{tab:processed-obs}Total number and weighted sum of records processed.") |>
  kable_styling(full_width = FALSE, font_size = 8) |>
  add_header_above(c(" " = 1, "Main" = 2, "Episode" = 2 )) %>%
  column_spec(3, border_right = TRUE)

table_df_processed
```

Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the TUS PUMF 2015 Main File (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing. Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the TUS PUMF 2015 _Episode file_ (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}. Only the unique active travel episode appear in Table \ref{tab:ep-2015-processed} since the records were filtered to select cases with walking or cycling episodes. For both cases, Tables \ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain labeled variables, facilitating the interpretation of the data.

```{r gss-processed-file-2015, echo=FALSE, cache=FALSE}
main_2015_processed <- gss_main_2015[1:10, 1:6]  |>
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the 2015 TUS Main File.") |>
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) |>
  footnote(general = c("Legend: `PUMFID`: record identification. `WGHT_PER`:  person weight. `SURVMNTH`: survey month of data collection. `AGEGR10`: age group of the respondent. `SEX`: sex of the respondent. `MARSTAT`: marital status of the respondent."), threeparttable = T)

main_2015_processed

ep_2015_unprocessed <- gss_episodes_renamed |>
  filter(PUMFID == 10041 & YEAR == 2015) |>
  dplyr::select(PUMFID, WGHT_EPI, act_label, DURATION, orig_label, dest_label, MODE) |>
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    caption = "\\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number `10041` of the 2015 GSS survey.", 
       col.names = c("PUMFID", "WGHT_EPI","Activity","Duration","Origin", "Destination", "Mode")) |> 
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) |>
  footnote(general = c("Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT_EPI: episode's weight. TUI_01: activity code. DURATION: episode's duration. LOCATION: episode's location."), threeparttable = T)

ep_2015_unprocessed
```

## Descriptive statistics

Considering all TUS analyzed, we identified `r nrow(gss_episodes)` episodes that recorded active travel episodes, with trip duration ranging from 0 to 900 minutes, to twelve different destinations. {ActiveCA} includes all these episodes ready for analysis. Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2022, with measures of trip duration in minutes. The 1986 survey did not include bicycle trips.

```{r table-01, echo=FALSE, cache=FALSE}
# Calculate min, mean, and max statistic (duration)
stats_data <- gss_episodes_renamed |>
  group_by(YEAR, MODE) |>
  dplyr::summarise(
    min = min(DURATION),
    mean = sum(DURATION * WGHT_EPI)/sum(WGHT_EPI),
    median =  Hmisc::wtd.quantile(DURATION, weights = WGHT_EPI, probs = 0.5),
    max = max(DURATION),
    std = sd(DURATION),
    .groups = "drop"
  ) |>
  pivot_longer(cols = c(min, mean, median, max, std), names_to = "Statistic", values_to = "Value") |>
  pivot_wider(names_from = YEAR, values_from = Value)

# Combine counts and stats into one data frame
stats_data <- stats_data |>
  mutate(MODE = factor(MODE, levels = c("Walking", "Cycling"))) |>
  arrange(MODE, Statistic) |>
  mutate(across(where(is.numeric), ~ round(.x, 0)))

# Rename the 'MODE' column to 'Mode' before generating the table
stats_data <- stats_data |>
  rename(Mode = MODE) |>
  dplyr::select(Mode, Statistic, everything()) |> 
   mutate(Statistic = case_when(
    Statistic == "max" ~ "Maximum",
    Statistic == "min" ~ "Minimum",
    Statistic == "mean" ~ "Mean",
    Statistic == "median" ~ "Median",
    Statistic == "std" ~ "Standard deviation",
    TRUE ~ Statistic))

# Creating a latex table
styled_ch03_table_01 <- stats_data |>
  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
        caption = "\\label{tab:table-01}Descriptive statistics for episodes with active transport records",  
        align = c('l', 'l', rep('c', ncol(stats_data) - 2))) |>
  kable_styling(full_width = FALSE, font_size = 8) |>
  column_spec(1, bold = TRUE) |>
  add_header_above(c(" " = 2, "Year" = ncol(stats_data) - 2)) |>
  collapse_rows(columns = 1, valign = "top", latex_hline = "major")

# Print the table
styled_ch03_table_01
```

Table \ref{tab:table-01} shows that, until 2022 the median values for walking trips were 10 minutes, increasing to 15 minutes in the last survey. In the case of cycling trips, the duration fluctuated over the years, ranging from 10 to 30 minutes. The table also highlights very high maximum values, particularly for walking trips, with recorded episodes exceeding 4 hours in all cases.

{ActiveCA} also enables visual analysis of active travel in Canada using exploratory data analysis techniques. Figure \ref{fig:figure-01} shows walking trips from 2022 through heat maps. This graph uses color gradients to represent the percentage of trips between various origins and destinations, with darker colors indicating higher percentages and lighter colors representing less frequent routes. For conciseness, we omitted the heat maps for the other years analyzed.

In 2022, `home` served as a central hub for most trips, with fewer than 10% of journeys not involving it as either a starting point or destination. The most common trip types were from `home` to `work or school` (17%) and the reverse, from `work or school` to `home` (13%). Notably, 7% of trips began and ended at `home`, often reflecting leisure activities such as short walks or dog walking. `Grocery stores` were also a key destination, comprising 10% of trips departing from `home.`

```{r heatmaps, echo=FALSE, cache=FALSE, include=FALSE}
# Walking heat map
walking_hm_fig <- gss_impedances_renamed %>%
  filter(YEAR == 2022, MODE == 'Walking') %>%
  group_by(orig_label, dest_label, YEAR) %>%
  dplyr::summarise(n = sum(WGHT_EPI), .groups = "drop") %>%
  group_by(YEAR) %>%
  mutate(percentage = n / sum(n) * 100) %>%
  ungroup() %>%
  ggplot(aes(x = dest_label,
    y = orig_label)) +
    geom_tile(aes(fill = percentage)) +
    labs(x = "Destination",
    y = "Origin",
    fill = "Percentage (%)") + 
    #geom_text(aes(label = round(percentage)), color = "black", size = 3) + 
    theme_bw() + 
    scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd")) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Saving figure
ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/walking_hm_2022.jpg"), 
       plot = walking_hm_fig,
       width = 11, 
       height = 7, 
       units = "cm", 
       dpi = 300)
```

```{r figure-01, echo=FALSE, out.width="100%", fig.cap="Percentage of walking trips categorized by origin and destination", fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/walking_hm_2022.jpg"))
```

The {ActiveCA} dataset also includes information on the type of population centre in which respondents reside - specifically, whether they live in a CMA, a CA, or outside these areas - as well as the respondent’s province. This information is important, as patterns of active travel often differ between metropolitan and non-metropolitan populations. For example, Table \ref{tab:cma-durations} presents the median walking durations by population centre type and province for 2022. Overall, respondents living in CMA/CA areas tend to report higher median walking durations compared to those living outside these centres. The most pronounced difference is observed in Nova Scotia: metropolitan residents reported a median walking duration of 30 minutes, whereas non-metropolitan residents reported a median of only 5 minutes.

```{r cma-durations, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}
median_duration_cma <- gss_episodes_renamed %>%
  filter(YEAR == 2022, 
         MODE == 'Walking') %>%
  group_by(Province, MODE, Pop_centre) %>%
  summarise(
    Median = Hmisc::wtd.quantile(DURATION, weights = WGHT_EPI, probs = 0.5),
    .groups = "drop") %>%
  select(-MODE) %>% 
  pivot_wider(
    names_from = Pop_centre,
    values_from = Median) %>% 
  mutate(Province = as.character(Province)) %>%
  arrange(Province)

# Latex table
median_duration_table <- median_duration_cma |>
  kable(
    format = "latex", 
    booktabs = TRUE, 
    caption = "\\label{tab:cma-durations}Differences in walking duration between provinces and population centre type.") |>
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) |>
  add_header_above(c(" " = 1, "Population centre type" = 2)) |> 
  footnote(general = c("'CMA' denotes Census Metropolitan Area and 'CA' denotes Census agglomeration."), threeparttable = T)

median_duration_table 
```

The package also enables obtaining insights from the main processed files. Figure \ref{fig:figure-stress} present how the level of stress varied among respondents depending on their marital status in 2022. According to this plot, married respondents reported the highest level of stress, relating to feel stressed every day, with 15% of possible cases.

```{r main_stress_figure, include = FALSE}
main_stress_figure <- gss_main_2022 |>
  filter(!is.na(MARSTAT), !is.na(GTU_110)) |>
  group_by(MARSTAT, GTU_110) |>
  dplyr::summarise(n = sum(WGHT_PER), .groups = "drop") |>
  mutate(percentage = n / sum(n) * 100) |>
  ungroup() |>
  ggplot(aes(x = GTU_110,
    y = MARSTAT)) +
    geom_tile(aes(fill = percentage)) +
    labs(x = "Marital status of the respondent",
    y = "How often do you feel rushed?",
    fill = "Percentage (%)") + 
    geom_text(aes(label = round(percentage)), color = "black", size = 3) + 
    theme_bw() + 
    scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlGn")) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Save the combined plot to a file
ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/main_stress_figure.jpg"), 
       plot = main_stress_figure,
       width = 15, 
       height = 10, 
       units = "cm", 
       dpi = 300)
```

```{r figure-stress, echo=FALSE, out.width="100%", fig.cap=paste0("Level of stress among respondents of different marital statuses (2015)."), fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/main_stress_figure.jpg"))
```

# Python integration

{ActiveCA} also provides a Jupyter Notebook containing a Python script that demonstrates how to read R data files (.rda) and convert them into Pandas DataFrames. This process allows users to work with and utilize the datasets available in {ActiveCA} within a Python project.

# Concluding remarks

This paper presents {ActiveCA}, an open data product that provides analysis-ready 
data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) 
of TUS GSSs on active travel in Canada. In the form of an `R` data package, 
{ActiveCA} was developed after collecting, cleaning, and processing the survey data, 
providing information on origins, destinations, and duration of active travel, as
well other information.

Although we did not select non-AT episodes, the process for obtaining them is very similar to that used for selecting AT episodes. Researchers interested in non-AT modes can use our framework to guide their methodology, making the small but necessary adjustments. We focused exclusively on AT episodes because the {ActiveCA} package is part of a larger project aimed at analyzing the historical evolution of active travel behaviour in Canada.

The value of {ActiveCA} lies in its transparency, accessibility, and ease of use,
which facilitates the addition of complementary data sets in the future. `R` users
can seamlessly explore TUS walking and cycling episodes, 
with the option to suggest enhancements to the package as needed. This article adopts
the structure proposed by Anastasia and Páez [-@soukhov2023], whose work provided
essential guidance for the creation of this package. Similarly, we aim to contribute
to the academic community by promoting transparent research practices that encourage
replication and innovation in related fields. We believe that {ActiveCA} will serve
as a basis for further research on TUS and for the integration of additional data by
the authors or the wider open source community.

# Declaration of Conflicting Interests

The author(s) declared no potential conflicts of interest with respect
to the research, authorship, and/or publication of this article.

# Funding

The author(s) disclosed receipt of the following financial support for
the research, authorship, and/or publication of this article: This work
was supported by the Social Sciences and Humanities Research Council of
Canada (*More description about the funding source after the review
process*).

# ORCID

Author 1

Author 2

Author 3

# Data availability statement

The {ActiveCA} R data package can be found and installed on Github
(*link*).

For review purposes, the package is currently available as a tar.gz file that can 
be installed by R. The file can be obtained from this anonymous location:

https://user.fm/files/v2-3d261d0b2aa47fcad096cd9e49fd5cf8/ActiveCA.zip

# References {#references .unnumbered}