### 2025-06-01 paper/Manuscript-Data-Package-1Round-Review.Rmd
Create a document with all manuscript changes and add answer letters for the reviewers.

---
title: "ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel"
author:
  - name: "Anon1"
    email: "Anon1@anon.org"
    affiliation: CODE
  - name: "Anon2"
    email: "Anon2@anon.org"
    affiliation: CODE
  - name: "Anon3"
    email: "Anon3@anon.org"
    affiliation: CODE
address:
  - code: CODE
    address: "Some school"
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
  "This paper describes {ActiveCA}, an open data product with Canadian time use data. {ActiveCA} is an R data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) cicles from the General Social Survey (GSS). Active travel is characterized by mode, with walking being part of every cycle and bicycling starting in 1992. The attributes of active trips are the types of locations of origins and destinations, the duration of trips, and episode weights for expanding the trips to population-wide estimates. Based on the year of the survey, a variety of locations are coded. In earlier cycles, these include home, work or school, and other's home, whereas in later cycles these are augmented with locations such as grocery stores, restaurants, outdoor destinations, and others. The geographical resolution includes the province and whether the episode was in an urban or rural setting."

keywords: Active; mobility; walking; cycling; travel time; time-use; General Social Survey 
classoption:
  - Royal
  - times
bibliography: bibfile.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  warning = FALSE,
  comment = '', 
  out.width = "1\\linewidth")

knitr::knit_hooks$set(inline = function(x) {
  if (is.numeric(x)) {
    format(x, big.mark = ",", scientific = FALSE)
  } else {
    x
  }
})

opts <- options(knitr.kable.NA = "")
```

```{r install-data-package, include=FALSE}
if (!require("ActiveCA", character.only = TRUE)) {
      remotes::install_github("dias-bruno/ActiveCA",
                        build_vignettes = TRUE)
}

rm(list=ls())
```

```{r load-packages, include=FALSE, cache=FALSE}
# Load packages
library(ActiveCA)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(reshape2)
library(rlang)
library(scales)
library(gridExtra)
library(ggpubr)
library(grid)
library(stats)
library(Hmisc)
```

```{r load-data, include = FALSE}
# Load data
data("gss_episodes")
data("gss_impedances")

# Original main and _Episode files_ from 2015 to display table examples
gss_m_2015 <- read.csv(paste0(here(),
                              "/data-raw/source-files/Time-use-2015/gss-m.csv"))
gss_e_2015 <- read.csv(paste0(here(), 
                              "/data-raw/source-files/Time-use-2015/gss-e.csv"))

# Processed _Main files_
data("gss_main_2022")
data("gss_main_2015")
data("gss_main_2010")
data("gss_main_2005")
data("gss_main_1998")
data("gss_main_1992")
data("gss_main_1986")
```

```{r rename-labels, include = FALSE}
# Renaming 'dest_label' and 'orig_label' values to reduce table space
gss_episodes_renamed <- gss_episodes |>
  mutate(dest_label = case_when(
    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
    dest_label == "Restaurant, bar or club" ~ "Restaurant",
    dest_label == "Sports centre, field or arena" ~ "Sport area",
    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
    dest_label == "In the neighbourhood" ~ "Neighbourhood",
    dest_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ dest_label  # Keep all other values as they are
  ), 
  orig_label = case_when(
    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
    orig_label == "Restaurant, bar or club" ~ "Restaurant",
    orig_label == "Sports centre, field or arena" ~ "Sport area",
    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
    orig_label == "In the neighbourhood" ~ "Neighbourhood",
    orig_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ orig_label))

gss_impedances_renamed <- gss_impedances |>
  mutate(dest_label = case_when(
    dest_label == "Grocery store, other stores or mall" ~ "Grocery store",
    dest_label == "Restaurant, bar or club" ~ "Restaurant",
    dest_label == "Sports centre, field or arena" ~ "Sport area",
    dest_label == "Medical, dental or other health clinic" ~ "Health clinic",
    dest_label == "In the neighbourhood" ~ "Neighbourhood",
    dest_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ dest_label  # Keep all other values as they are
  ), 
  orig_label = case_when(
    orig_label == "Grocery store, other stores or mall" ~ "Grocery store",
    orig_label == "Restaurant, bar or club" ~ "Restaurant",
    orig_label == "Sports centre, field or arena" ~ "Sport area",
    orig_label == "Medical, dental or other health clinic" ~ "Health clinic",
    orig_label == "In the neighbourhood" ~ "Neighbourhood",
    orig_label == "Library, museum or theatre" ~ "Cultural venues",
    TRUE ~ orig_label))
```

# Introduction

The objective of this paper is to introduce {ActiveCA}, an open data
product with data from all Time Use Survey (TUS) cycles of the General
Social Surveys (GSSs). Open data products (ODPs) are the outcome of a
process that transforms raw data (open or not) into analysis-ready data,
following a transparent process in which all stages of development
follow open principles [@arribas-bel2021]. ODPs, while still open,
differ from general open data in their degree of ease of access, their
heightened usability, and potentially the value they add to the raw
data.

{ActiveCA} provides analysis-ready data concerning active travel in
Canada spanning a period of almost 40 years, obtained from the TUS
cycles of the Canadian GSS. The GSS program is designed to provide
cross-sectional data on topics of interest to improve the well-being of
Canadians. As part of this program, every five to seven years the survey
is done on the topic of time use (TUS). Concretely, {ActiveCA} covers
Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015),
and 34 (2022) of the TUS. Time use data in these surveys is coded using
a very fine grain, from time spent in chores, leisure, and sleeping, to
time spent working or at school. These surveys have proved valuable in
investigations of mobility and quality of life [@spinneyTransport2009],
the relationship between active travel and transit use
[@lachapelleLonger2016], and travel behavior and time poverty
[@kimFacing2024], to name but a few examples.

Using Public Use Microdata Files (PUMFs) from the Time Use Survey (TUS),
we extracted all data necessary to characterize active travel in
Canada - specifically, episodes in which the activity involved moving
between an origin and a destination by walking or cycling. Although
Statistics Canada provides PUMFs and accompanying documentation for the
GSS program [see @statisticscanada2024], accessing and preparing these
files for analysis is not a straightforward task due to their size and
complexity. The process of extracting information of interest from the
source files is time-consuming, tedious, and challenging and/or prone to
error due to the expertise required to work with these files. To create
{ActiveCA} we selected, labelled, and processed the TUS cycles to make
them ready for analysis.

{ActiveCA} is distributed as an `R` package with a number of data objects
and their documentation. `R` packages contain code, data, and
documentation in a standardized format that can be installed by R users
via a software repository, such as CRAN (Comprehensive R Archive
Network) or GitHub, which makes them an adroit medium to distribute
analysis-ready data.

Given the level of interest in active travel [e.g.,
@mccurdySupport2023], reducing the barriers to using data contained in
rich, but difficult to access and use surveys, such as TUS, is a worthy
endeavour that can only improve data-driven decisions in transportation,
urban, and health policy.

The rest of this paper discusses the sources of data, and the process
implemented to retrieve and package them. Then, we explain how to use
the package and show some selected examples of analysis to whet the
imagination of potential users. This ODP provides not only data that are
easy to use, but also all the code and documentation that make this a
reproducible research project. In summary, {ActiveCA} aims to implement
and inspire the best principles of open spatial sciences
[@paez_open_2021; @brunsdon_opening_2021].

# The Time Use Survey (TUS) collection

Statistics Canada [-@statisticscanada2024] conducts GSS surveys to
obtain data on social trends to track changes in Canadians' living
conditions and well-being over time. The series of survey on time use
are used to understand how Canadian residents spend and manage their
time, and what factors contribute to their happiness and stress. The GSS
program was created in 1985, and is serialized to provide a collection
of annual, representative cross-sectional surveys.

The topics of the survey cycle every few years to cover topics that
include family, health, social identity, and every five to seven years
time use. The first Canadian TUS done as part of the GSS program was
conducted in 1986, and the most recent was completed in 2022. These Time
Use Surveys [@statisticscanada2022] collect data on respondents'
participation and time spent on a wide range of everyday activities
using a 24-hour retrospective diary, with information on the location of
these activities (e.g. at home, at work, etc.) and, for non-personal
activities, the people who were present with the respondent at the time
of the activity. In addition, time-use surveys also cover topics related
to leisure time, work-life balance, health, commuting, culture and
sports, and many others.

TUS allows researchers to identify the origin and destination of trips,
travel times and modes of transport used, providing a valuable dataset
for analyzing active travel behavior. It also provides the empirical
basis for tools used in transportation analysis, such as the development
of impedance functions for accessibility analysis, a measure of the ease
with which people can reach destinations and opportunities
[@hansen1959]. The Canadian TUS is unique at the national level in
collecting detailed information on travel behaviour. Its consistent
application across survey cycles enables the identification of long-term
trends, with some questions present in the questionnaires since the
first application of the survey.

Most respondents to the 2022 TUS completed it online, reflecting
Statistics Canada's effort to adapt to technological changes and growing
time demands by offering greater flexibility and convenience
[@statisticscanada2022]. While such methodological changes may affect
data comparability over time, it is not possible to determine whether
observed differences result from actual population changes or shifts in
data collection methods. Despite rigorous efforts to ensure data
quality, the use of electronic questionnaires may have influenced
estimates. Statistics Canada assessed the impact of collection mode on a
limited set of key questions, constrained by sample size. Importantly,
none of the variables used in this research are `R` package in the 2022
PUMF User Guide as unsuitable for trend analysis.

Until 2022, Statistics Canada employed a telephone-based sampling frame,
which was replaced by a dwelling-based frame in the most recent cycle.
Each survey cycle spans a 12-month period, typically from July to the
following July. The target population includes all Canadians aged 15 and
over, excluding residents of the Yukon, Northwest Territories, and
Nunavut, full-time institutional residents, and individuals living on
Indigenous reserves.

The survey encompasses both rural and urban areas, including
metropolitan and non-metropolitan regions, to ensure a diverse and
representative sample. For sampling, the ten provinces were divided into
geographic strata. Several Census Metropolitan Areas (CMAs) - such as
St. John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa,
Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and
Vancouver - were treated as separate strata. Additional strata grouped
other CMAs within Quebec, Ontario, and British Columbia, as well as
non-CMA areas within each province.

The Public Use Microdata are released by Statistics Canada in two files:
a *Main file* and an *Episode file*. The files are linked by keys that
identify households, individuals, and episodes (i.e., activities)
conducted by individuals. We discuss these files in more detail in the
following section.

## The Main file

The *Main file* of the time use survey compiles a large array of
aggregated data, summarizing the answers to the questionnaire that
describe households and individuals, as well as derived variables that
summarize the respondents' use of time use across different activities,
locations, and social interactions. This file documents the time and
duration that respondents allocate to each activity and location. The
*Main file* provides a overview of daily routines and social dynamics,
not focusing on individual activity episodes. Additionally, this file
categorizes activities into bigger groups and subcategories,
facilitating the data's analytical utility with additional metrics such
as total transit time, time spent with household members, and counts of
activities and episodes.

Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first
six variables of the TUS PUMF 2015 *Main file* (Cycle 29). Each row in
the table correspond to a survey respondent, while the columns refer the
following information: record identification (`PUMFID`), the person's
weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`),
the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and
the respondent's marital status (`MARSTAT`).

The *Main file* of the 2015 GSS surveys includes a total of 17,390
respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and
848 variables. For discrete variables, Statistics Canada has assigned
specific codes to the possible values, with each code accompanied by a
label. For instance, in the case of the variable `SURVMNTH`, a value of `1`
means `January 2016`, `2` means `February 2016`, `3` corresponds to `March 2016`,
and so on.

As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not
labeled. Additionally, the format of the tables (comma-separated values)
does not allow for the specification of variable types (whether a
variable is continuous or discrete), which can lead to mistakes analysts
who have limited experience working with PUMFs.

```{r gss-main-file-2015}
main_2015_unprocessed <- gss_m_2015[1:10, 1:6]  |>
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main file of the 2015 GSS.") |>
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) |>
  footnote(general = c("Legend: PUMFID: record identification. WGHT_PER:  person weight. SURVMNTH: survey month of data collection. AGEGR10: age group of the respondent. SEX: sex of the respondent. MARSTAT: marital status of the respondent."), threeparttable = T)

main_2015_unprocessed
```

## The Episode file

The episode is a much bigger file that records detailed data for each
activity episode reported by respondents. Each episode represents a
single activity and its duration, and the sum of all episodes throughout
the day adds up to 24 hours. Each entry in this file includes the start
and end times of the activity, the duration, location, and accompanying
social context, informing when and where activities occurred and with
whom. The focus of the *Episode file* is not on the characteristics of
the respondents, but on the characteristics of the activities, and the
data are structured around the numerous activity instances that compose
a day of the respondent. Although respondent-specific characteristics
are not included within the *Episode file*, it is possible to link the
*Main file* and the *Episode file* by using a key present in both the
*Main* and *Episode files*.

Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an
example of the *Main file* structure, Table
\ref{tab:ep-2015-unprocessed} presents the first seven episodes for the
record identification number 10041 and some variables from the TUS PUMF
2015 *Episode file* (Cycle 29). Each row in the table corresponds to an
episode associated with the specified record identification (`PUMFID` =
`10041`), including the episode's weight (`WGHT_EPI`), episode number
(`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and
episode location (`LOCATION`).

In total, the *Episode file* of the 2015 GSS surveys contains 274,108
records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527
variables detailing the episodes. Similar to the *Main file*, Statistics
Canada has created codes for the discrete variables, with each value
corresponding to a label.

In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this
respondent began the diary description by sleeping at home (`TUI_01 = 1`
and `LOCATION = 300`) for 210 minutes, followed by 40 minutes of personal
hygiene (`TUI_01 = 2`). The respondent then spent 15 minutes on personal
care activities, such as getting ready for school, supervising homework,
reading, playing, reprimanding, or providing educational or emotional
support, as indicated by `TUI_01 = 27.` Next, they recorded a travel
episode, walking for 15 minutes (`TUI_01 = 7` and `LOCATION = 315`), where
both the origin and destination were their home. Such trips, where the
journey starts and finishes at home, can be classified as recreational
or leisure trips. Next, the respondent spent 3 hours searching for a job
(`TUI_01 = 9`), took a 15-minute lunch break (`TUI_01 = 6`), and then
cleaned the house (`TUI_01 = 18`) for two hours. Table
\ref{tab:ep-2015-unprocessed} displays only six variables out of the 527
available. As shown, since the dataset does not label the variable
values, decoding them can be both time-consuming and challenging.

```{r gss-epi-file-2015}
ep_2015_unprocessed <- gss_e_2015 |>
  filter(PUMFID == 10041 & EPINO <= 7) |>
  dplyr::select(PUMFID, EPINO,WGHT_EPI,TUI_01, DURATION, LOCATION) |>
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
    caption = "\\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number 10041.") |>
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) |>
  footnote(general = c("Legend: PUMFID: record identification. EPINO: episode number. WGHT_EPI: episode's weight. TUI_01: activity code. DURATION: episode's duration. LOCATION: episode's location."), threeparttable = T)

ep_2015_unprocessed
```

# Data process

Figure \ref{fig:process-figure} presents a diagram illustrating the
processes applied to the *Main* and *Episode files* to create the
{ActiveCA} datasets. For each TUS cycle, we reviewed the *Episode files*
to identify movement episodes involving walking or cycling. This allowed
us to also select the activities immediately before and after the
movement episode, helping to infer the trip’s purpose and determine its
origin and destination. Active trips were identified by their
corresponding activity codes, accounting for variations across survey
cycles.

Next, we labeled the coded variables with their appropriate
descriptions, classifying each origin and destination, mode of travel,
and time spent in the active trip. Categorical variables were converted
into factor variables, and ordinal variables were defined as ordered
factors.

For the *Main files*, we selected socioeconomic variables to help
profile individuals engaged in active travel episodes. These included
key indicators such as age group, sex, marital status, and number of
children, among others. As with the *Episode files*, we labeled and
factored the coded variables. Using the appropriate identifiers, we
joined the Episode and *Main files* to determine the province and the
type of population centre - whether it is a CMA, a Census Agglomeration
(CA), or a non-CMA/CA area.

As output, the {ActiveCA} package provides processed datasets of walking
and cycling episodes by year, a unified dataset that contains all
walking and cycling episodes across TUS cycles, and processed main
files (not unified, as socioeconomic and demographic variables and
their categories differ across TUS cycles).

```{r process-figure, echo=FALSE, out.width="100%", fig.cap="Diagram with the processes applied to the main (blue arrows) and episode files (green arrows) to obtain the ActiveCA datasets.", fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/RPackages - ActiveCA.jpeg"))
```

# {ActiveCA} data sets

This section presents some potential applications of the {ActiveCA} `R` package. In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses. The installation instruction and also some examples of application of the {ActiveCA} `R` package are available in the `vignettes`, available in the Github repository.

## Active episodes

```{r processed-files, include=FALSE}
df_main_processed <- data.frame(
  'Survey' = c("Main", "Main","Main","Main","Main","Main","Main"),
  'Year' = c(2022, 2015, 2010, 2005, 1998, 1992, 1986),
  'Count' = c(nrow(gss_main_2022),
              nrow(gss_main_2015),
              nrow(gss_main_2010),
              nrow(gss_main_2005),
              nrow(gss_main_1998),
              nrow(gss_main_1992),
              nrow(gss_main_1986)),
  'Weighted' = c(sum(gss_main_2022$WGHT_PER),
              sum(gss_main_2015$WGHT_PER),
              sum(gss_main_2010$WGHT_PER),
              sum(gss_main_2005$WGHT_PER),
              sum(gss_main_1998$WGHTFIN),
              sum(gss_main_1992$FWGHT),
              sum(gss_main_1986$FWGT_OS/10000)))

df_ep_processed <- gss_episodes |> 
  group_by(YEAR) |>
  summarise(Survey = 'Episode', 
            Count = n(), 
            Weighted = sum(WGHT_EPI)) |>
  rename(Year = YEAR) |>
  dplyr::select(Survey, Year, Count, Weighted) |> 
  arrange(-Year)

df_processed <- rbind(df_main_processed,df_ep_processed)

# For table

df_ep_processed_table <- gss_episodes |> 
  group_by(YEAR) |>
  summarise(Survey_ep = 'Episode', 
            Count_ep = n(), 
            Weighted_ep = sum(WGHT_EPI)) |>
  rename(Year_ep = YEAR) |>
  dplyr::select(Survey_ep, Year_ep, Count_ep, Weighted_ep) |> 
  arrange(-Year_ep)

df_processed_table <- cbind(df_main_processed,df_ep_processed_table)
```

Table \ref{tab:processed-obs} displays the total number of records
processed for Main and *Episode files*. For the *Main files*, a total of
`r sum(df_main_processed$Count)` records were processed, referring to
all records from the Time Use Surveys from 1986 to 2022, that together
represents more of `r sum(df_main_processed$Weighted)` respondents. It
also presents the total cases of active trips episodes identified. In
total `r sum(df_ep_processed$Count)` records with register of active
travel activity. Together, these records account for
`r sum(df_ep_processed$Weighted)` episodes.

```{r table_df_processed}

table_df_processed <- df_processed_table |> 
  select(-c(Survey_ep, Survey, Year_ep)) %>% 
  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
        caption = "\\label{tab:processed-obs}Total number and weighted sum of records processed.") |>
  kable_styling(full_width = FALSE, font_size = 8) |>
  add_header_above(c(" " = 1, "Main" = 2, "Episode" = 2 )) %>%
  column_spec(3, border_right = TRUE)

table_df_processed
```

Table \ref{tab:main-2015-processed} shows the first ten rows and first
six variables of the TUS PUMF 2015 Main File (Cycle 29), displayed in
\ref{tab:main-2015-unprocessed} before our processing. Table
\ref{tab:ep-2015-processed} presents the walking episodes for the record
identification number `10041` from the TUS PUMF 2015 *Episode file* (Cycle
29), previously displayed in Table \ref{tab:ep-2015-unprocessed}. Only
the unique active travel episode appear in Table
\ref{tab:ep-2015-processed} since the records were filtered to select
cases with walking or cycling episodes. For both cases, Tables
\ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain
labeled variables, facilitating the interpretation of the data.

```{r gss-processed-file-2015, echo=FALSE, cache=FALSE}
main_2015_processed <- gss_main_2015[1:10, 1:6]  |>
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the 2015 TUS Main File.") |>
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) |>
  footnote(general = c("Legend: PUMFID: record identification. WGHT_PER:  person weight. SURVMNTH: survey month of data collection. AGEGR10: age group of the respondent. SEX: sex of the respondent. MARSTAT: marital status of the respondent."), threeparttable = T)

main_2015_processed

ep_2015_unprocessed <- gss_episodes_renamed |>
  filter(PUMFID == 10041 & YEAR == 2015) |>
  dplyr::select(PUMFID, WGHT_EPI, act_label, DURATION, orig_label, dest_label, MODE) |>
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    caption = "\\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number 10041 of the 2015 GSS survey.", 
       col.names = c("PUMFID", "WGHT_EPI","Activity","Duration","Origin", "Destination", "Mode")) |> 
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) |>
  footnote(general = c("Legend: PUMFID: record identification. EPINO: episode number. WGHT_EPI: episode's weight. TUI_01: activity code. DURATION: episode's duration. LOCATION: episode's location."), threeparttable = T)

ep_2015_unprocessed
```

## Descriptive statistics

Considering all TUS analyzed, we identified `r nrow(gss_episodes)`
episodes that recorded active travel episodes, with trip duration
ranging from 0 to 900 minutes, to twelve different destinations.
{ActiveCA} includes all these episodes ready for analysis. Table
\ref{tab:table-01} presents descriptive statistics on walking and
cycling trips between 1986 and 2022, with measures of trip duration in
minutes. The 1986 survey did not include bicycle trips.

```{r table-01, echo=FALSE, cache=FALSE}
# Calculate min, mean, and max statistic (duration)
stats_data <- gss_episodes_renamed |>
  group_by(YEAR, MODE) |>
  dplyr::summarise(
    min = min(DURATION),
    mean = sum(DURATION * WGHT_EPI)/sum(WGHT_EPI),
    median =  Hmisc::wtd.quantile(DURATION, weights = WGHT_EPI, probs = 0.5),
    max = max(DURATION),
    std = sd(DURATION),
    .groups = "drop"
  ) |>
  pivot_longer(cols = c(min, mean, median, max, std), names_to = "Statistic", values_to = "Value") |>
  pivot_wider(names_from = YEAR, values_from = Value)

# Combine counts and stats into one data frame
stats_data <- stats_data |>
  mutate(MODE = factor(MODE, levels = c("Walking", "Cycling"))) |>
  arrange(MODE, Statistic) |>
  mutate(across(where(is.numeric), ~ round(.x, 0)))

# Rename the 'MODE' column to 'Mode' before generating the table
stats_data <- stats_data |>
  rename(Mode = MODE) |>
  dplyr::select(Mode, Statistic, everything()) |> 
   mutate(Statistic = case_when(
    Statistic == "max" ~ "Maximum",
    Statistic == "min" ~ "Minimum",
    Statistic == "mean" ~ "Mean",
    Statistic == "median" ~ "Median",
    Statistic == "std" ~ "Standard deviation",
    TRUE ~ Statistic))

# Creating a latex table
styled_ch03_table_01 <- stats_data |>
  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
        caption = "\\label{tab:table-01}Descriptive statistics for episodes with active transport records",  
        align = c('l', 'l', rep('c', ncol(stats_data) - 2))) |>
  kable_styling(full_width = FALSE, font_size = 8) |>
  column_spec(1, bold = TRUE) |>
  add_header_above(c(" " = 2, "Year" = ncol(stats_data) - 2)) |>
  collapse_rows(columns = 1, valign = "top", latex_hline = "major")

# Print the table
styled_ch03_table_01
```

Table \ref{tab:table-01} shows that, until 2022 the median values for
walking trips were 10 minutes, increasing to 15 minutes in the last
survey. In the case of cycling trips, the duration fluctuated over the
years, ranging from 10 to 30 minutes. The table also highlights very
high maximum values, particularly for walking trips, with recorded
episodes exceeding 4 hours in all cases.

{ActiveCA} also enables visual analysis of active travel in Canada using
exploratory data analysis techniques. Figure \ref{fig:figure-01} shows
walking trips from 2022 through heat maps. This graph uses color
gradients to represent the percentage of trips between various origins
and destinations, with darker colors indicating higher percentages and
lighter colors representing less frequent routes. For conciseness, we
omitted the heat maps for the other years analyzed.

In 2022, `home` served as a central hub for most trips, with fewer than
10% of journeys not involving it as either a starting point or
destination. The most common trip types were from `home` to `work or school`
(17%) and the reverse, from `work or school` to `home` (13%). Notably, 7% of
trips began and ended at `home`, often reflecting leisure activities such
as short walks or dog walking. `Grocery stores` were also a key
destination, comprising 10% of trips departing from `home.`

```{r heatmaps, echo=FALSE, cache=FALSE, include=FALSE}
# Walking heat map
walking_hm_fig <- gss_impedances_renamed %>%
  filter(YEAR == 2022, MODE == 'Walking') %>%
  group_by(orig_label, dest_label, YEAR) %>%
  dplyr::summarise(n = sum(WGHT_EPI), .groups = "drop") %>%
  group_by(YEAR) %>%
  mutate(percentage = n / sum(n) * 100) %>%
  ungroup() %>%
  ggplot(aes(x = dest_label,
    y = orig_label)) +
    geom_tile(aes(fill = percentage)) +
    labs(x = "Destination",
    y = "Origin",
    fill = "Percentage (%)") + 
    #geom_text(aes(label = round(percentage)), color = "black", size = 3) + 
    theme_bw() + 
    scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlOrRd")) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Saving figure
ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/walking_hm_2022.jpg"), 
       plot = walking_hm_fig,
       width = 11, 
       height = 7, 
       units = "cm", 
       dpi = 300)
```

```{r figure-01, echo=FALSE, out.width="100%", fig.cap="Percentage of walking trips categorized by origin and destination", fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/walking_hm_2022.jpg"))
```

The {ActiveCA} dataset also includes information on the type of
population centre in which respondents reside - specifically, whether
they live in a CMA, a CA, or outside these areas - as well as the
respondent’s province. This information is important, as patterns of
active travel often differ between metropolitan and non-metropolitan
populations. For example, Table \ref{tab:cma-durations} presents the
median walking durations by population centre type and province for
2022. Overall, respondents living in CMA/CA areas tend to report higher
median walking durations compared to those living outside these centres.
The most pronounced difference is observed in Nova Scotia: metropolitan
residents reported a median walking duration of 30 minutes, whereas
non-metropolitan residents reported a median of only 5 minutes.

```{r cma-durations, echo=FALSE, cache=FALSE, warning=FALSE, results="asis"}
median_duration_cma <- gss_episodes_renamed %>%
  filter(YEAR == 2022, 
         MODE == 'Walking') %>%
  group_by(Province, MODE, Pop_centre) %>%
  summarise(
    Median = Hmisc::wtd.quantile(DURATION, weights = WGHT_EPI, probs = 0.5),
    .groups = "drop") %>%
  select(-MODE) %>% 
  pivot_wider(
    names_from = Pop_centre,
    values_from = Median) %>% 
  mutate(Province = as.character(Province)) %>%
  arrange(Province)

# Latex table
median_duration_table <- median_duration_cma |>
  kable(
    format = "latex", 
    booktabs = TRUE, 
    caption = "\\label{tab:cma-durations}Differences in walking duration between provinces and population centre type.") |>
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) |>
  add_header_above(c(" " = 1, "Population centre type" = 2)) |> 
  footnote(general = c("CMA denotes Census Metropolitan Area and CA denotes Census agglomeration."), threeparttable = T)

median_duration_table 
```

The package also enables obtaining insights from the main processed
files. Figure \ref{fig:figure-stress} present how the level of stress
varied among respondents depending on their marital status in 2022.
According to this plot, married respondents reported the highest level
of stress, relating to feel stressed every day, with 15% of possible
cases.

```{r main_stress_figure, include = FALSE}
main_stress_figure <- gss_main_2022 |>
  filter(!is.na(MARSTAT), !is.na(GTU_110)) |>
  group_by(MARSTAT, GTU_110) |>
  dplyr::summarise(n = sum(WGHT_PER), .groups = "drop") |>
  mutate(percentage = n / sum(n) * 100) |>
  ungroup() |>
  ggplot(aes(x = GTU_110,
    y = MARSTAT)) +
    geom_tile(aes(fill = percentage)) +
    labs(x = "Marital status of the respondent",
    y = "How often do you feel rushed?",
    fill = "Percentage (%)") + 
    geom_text(aes(label = round(percentage)), color = "black", size = 3) + 
    theme_bw() + 
    scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "YlGn")) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Save the combined plot to a file
ggsave(file = paste0(here::here(),"/paper/Manuscript-figures/main_stress_figure.jpg"), 
       plot = main_stress_figure,
       width = 15, 
       height = 10, 
       units = "cm", 
       dpi = 300)
```

```{r figure-stress, echo=FALSE, out.width="100%", fig.cap=paste0("Level of stress among respondents of different marital statuses (2015)."), fig.align="center"}

knitr::include_graphics(paste0(here::here(),"/paper/Manuscript-figures/main_stress_figure.jpg"))
```

# Python integration

{ActiveCA} also provides a Jupyter Notebook containing a `Python` script
that demonstrates how to read `R` data files (.rda) and convert them into
Pandas DataFrames. This process allows users to work with and utilize
the datasets available in {ActiveCA} within a `Python` project.

# Concluding remarks

This paper presents {ActiveCA}, an open data product that provides
analysis-ready data from Cycles 2 (1986), 7 (1992), 12 (1998), 19
(2005), 24 (2010), 29 (2015), and 34 (2022) of TUS GSSs on active travel
in Canada. In the form of an `R` data package, {ActiveCA} was developed
after collecting, cleaning, and processing the survey data, providing
information on origins, destinations, and duration of active travel, as
well other information.

Although we did not select non-AT episodes, the process for obtaining
them is very similar to that used for selecting AT episodes. Researchers
interested in non-AT modes can use our framework to guide their
methodology, making the small but necessary adjustments. We focused
exclusively on AT episodes because the {ActiveCA} package is part of a
larger project aimed at analyzing the historical evolution of active
travel behaviour in Canada.

The value of {ActiveCA} lies in its transparency, accessibility, and
ease of use, which facilitates the addition of complementary data sets
in the future. R users can seamlessly explore TUS walking and cycling
episodes, with the option to suggest enhancements to the package as
needed. This article adopts the structure proposed by Anastasia and Páez
[-@soukhov2023], whose work provided essential guidance for the creation
of this package. Similarly, we aim to contribute to the academic
community by promoting transparent research practices that encourage
replication and innovation in related fields. We believe that {ActiveCA}
will serve as a basis for further research on TUS and for the
integration of additional data by the authors or the wider open source
community.

# Declaration of Conflicting Interests

The author(s) declared no potential conflicts of interest with respect
to the research, authorship, and/or publication of this article.

# Funding

The author(s) disclosed receipt of the following financial support for
the research, authorship, and/or publication of this article: This work
was supported by the Social Sciences and Humanities Research Council of
Canada (*More description about the funding source after the review
process*).

# ORCID

Author 1

Author 2

Author 3

# Data availability statement

The {ActiveCA} R data package can be found and installed on Github
(*link*).

For review purposes, the package is currently available as a tar.gz file
that can be installed by R. The file can be obtained from this anonymous
location:

<https://user.fm/files/v2-3d261d0b2aa47fcad096cd9e49fd5cf8/ActiveCA.zip>

# References {#references .unnumbered}