### 2025-08-14 README.Rmd
Final version for EPB publication

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ActiveCA

<!-- badges: start -->
<!-- badges: end -->

{ActiveCA} is an open data product in the form of an `R` data package with information about active travel in Canada. The data product is based on Public Use Microdata Files of Statistics Canada's [General Social Survey (GSS) program](https://www150.statcan.gc.ca/n1/pub/45-25-0001/index-eng.htm) with a focus on the [Time Use Survey](https://www150.statcan.gc.ca/n1/pub/45-25-0001/index-eng.htm#a7) cycles.

To reduce the barriers to the use of the Public Use Microdata Files, this data product has extracted all walking and cycling episodes and their corresponding episode weights. Origins and destinations are categorized, and it is possible to investigate active travel (number of episodes and their duration) for broad categories of purposes (e.g., travel to work or school, travel to other's home). {ActiveCA} also include some socioeconomic variables from the GSS Main Files to facilitate profiling the survey's respondents.

The data covers cycles Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 2022, spanning a period of more than thirty five years.

## Installation

You can install the development version of ActiveCA from [GitHub](https://github.com/dias-bruno/ActiveCA) with:

``` r
# install.packages("pak")
pak::pak("dias-bruno/ActiveCA")
```